Amphibian mast cells: barriers to deadly chytrid fungus infections

Version 1

Abstract

Abstract

Global amphibian declines are largely driven by deadly disease outbreaks caused by the chytrid fungus, Batrachochytrium dendrobatidis (Bd). In the time since these disease outbreaks were first discovered, much has been learned about the roles of amphibian skin-produced antimicrobial components and skin microbiomes in controlling Bd. Yet almost nothing is known about the roles of skin-resident immune cells in anti-Bd defenses. Notably, mammalian mast cells reside within and serve as key immune sentinels in barrier tissues like the skin. Thus, they are critical to immune recognition of pathogens and to orchestrating the ensuing immune responses. Accordingly, we investigated the roles of Xenopus laevis frog mast cells during Bd infections. Our findings indicate that enrichment of X. laevis skin mast cells confers significant anti-Bd protection and ameliorates the inflammation-associated skin damage caused by Bd infection. Moreover, enriching X. laevis mast cells promotes greater mucin content within cutaneous mucus glands and protects frogs from Bd-mediated changes to their skin microbiomes. Together, this work underlines the importance of amphibian skin-resident immune cells in anti-Bd defenses and introduces a novel approach for investigating amphibian host-chytrid pathogen interactions.

Introduction

Catastrophic declines of hundreds of amphibian species across six continents have been causally linked to the chytrid fungi, Batrachochytrium dendrobatidis (Bd) and Batrachochytrium salamandrivorans (Bsal)1, 2. Motile Bd zoospores readily colonize keratinized skin of adult amphibians and keratinized mouthparts of tadpoles3. Bd skin infections culminate in chytridiomycosis, ultimately disrupting the function of this respiratory and barrier tissue3, 4, 5. Effectively combatting chytrid infections requires a holistic understanding of amphibian cutaneous immune defenses. Research efforts up to this point have focused on antifungal capacities of amphibian skin-produced antimicrobial peptides6, 7; commensal antifungal products7, 8, 9; antifungal properties of mucus9, alkaloids10, and lysozymes11, 12 as well as the roles of antibodies4,13. Many of these studies suggest that Bd exposure can elicit some immune protection14. However, the contribution of skin-resident immune cells to amphibian anti-Bd defenses remains almost entirely unexplored.

Mammalian mast cells serve as sentinels of mucosal and connective tissues, concentrated in barrier tissues like skin, where they maintain homeostasis and regulate immune responses15. Other granulocyte-lineage cells such as neutrophils are generally not found in healthy tissues and only extravasate into sites of inflammation16. Consequently, mast cells are among the first immune cells to recognize and respond to skin-infiltrating pathogens. When activated, mast cells release pre-formed and de novo-generated immunomodulatory compounds that may serve to elicit, exacerbate, or ameliorate inflammatory responses15. Cells bearing hallmark mast cell cytology have been reported across a range of non-mammalian species17, 18, including amphibians19. Notably, the principal mast cell growth factor, stem cell factor (SCF, KIT ligand) required for mast cell differentiation and survival20 is expressed by all vertebrates examined to-date.

Here, we combine comprehensive in vitro and in vivo approaches to define the roles of amphibian (Xenopus laevis) mast cells during Bd infections. Our results provide compelling evidence that skin-resident immune cells contribute to anti-Bd defenses.

Results

Frog mast cells possess archetypal mast cell cytology and transcriptional profiles

We produced X. laevis recombinant (r)SCF, and used this reagent to generate mast cell cultures from bone marrow-derived myeloid precursors21. Mast cells were compared to bone marrow-derived neutrophilic granulocytes (hereafter referred to as ‘neutrophils’), differentiated using a recombinant X. laevis colony-stimulating factor-321 (rCSF3, i.e., granulocyte colony-stimulating factor; GCSF). While the neutrophil cultures were comprised of cells with hyper-segmented nuclei and neutral-staining cytoplasms (Fig. 1A), the mast cell cultures consisted predominantly of mononuclear cells with basophilic cytoplasm (Fig. 1B). We confirmed the granulocyte-lineage of X. laevis mast cells using specific esterase (SE) staining (Fig. 1D). As expected, X. laevis neutrophils were also SE-positive (Fig. 1C). Mast cell and neutrophil morphology was further explored with electron microscopy (Fig. 1E-H). SEM imaging demonstrated that X. laevis mast cells possess extensive folding of their plasma membranes (Fig. 1F). This mast cell-characteristic membrane ruffling appeared as projections resembling pseudopods via TEM, which further revealed electron-dense heterogenous granules, few mitochondria, and round to elongated nuclei (Fig. 1H) typical of mammalian mast cells22. X. laevis neutrophils also exhibited pronounced membrane ruffling (Fig. 1E) but strikingly distinct intracellular appearance including multilobed nuclei (Fig. 1G).

X. laevis bone marrow-derived mast cells possess classical mast cell cytology and transcriptional profiles.

Neutrophils (A, C, E, G) and mast cells (B, D, F, H) were stained with Giemsa (A, B) and Leder to visualize specific esterase activity (SE) (C, D) or imaged with scanning and transmission electron microscopy (SEM: E, F and TEM: G, H). (I) Heat map of the top 30 differentially expressed genes (DEGs) identified with RNA sequencing analyses of X. laevis mast cell (N=4) and neutrophil (N=4) cultures. Log2fold change in expression represented as color scale.

Frog mast cells and neutrophils also exhibited distinct transcriptional profiles of immune- related genes including those encoding lineage-specific transcription factors, immune receptors, downstream signaling components and adhesion molecules, as well as non-immune genes (Fig. 1I). Frog mast cells and neutrophils each expressed greater levels of lineage-specific transcription factors associated with mammalian mast cell (gata1, gata2, and mitf)23 and neutrophil (cebp family members)24 counterparts, respectively (Fig. 2A). Notably, mast cells expressed greater levels of enzyme and cytokine genes associated with tissue remodeling (carboxypeptidase; cpa25), immune suppression (indoleamine 2,3 dioxygenase-1; ido126) and amelioration of cutaneous inflammation (leukemia inhibitory factor, lif27; Fig. 2B). Conversely, neutrophils expressed predominantly proinflammatory enzymes and cytokine genes such as leukotriene 4 hydrolase (lta4h; Fig. 2B) and tumor necrosis factor alpha (tnfα, Fig. 2B). In addition, mast cells and neutrophils each had greater expression of genes encoding their respective growth factor receptors, kit and csf3r (Fig. 2B).

Frog mast cells and neutrophils possess gene profiles similar to their mammalian counterparts.

The differentially expressed genes from the RNA sequencing analyses of X. laevis mast cells and neutrophil cultures were profiled for those encoding (A) transcription factors associated with mast cell- or neutrophil-specific lineages and (B) granulocyte antimicrobial components and growth factor receptor genes. All depicted genes were significantly differentially expressed between the two populations, N=4 per group.

Enriching mast cells in frog skin offers protection against Bd

Although all granulocyte-lineage cells possess SE activity, mast cells are the predominant mononuclear granulocytes to reside in vertebrate tissues28. Therefore, we selectively enriched mast cells in X. laevis skin via subcutaneous rSCF administration (note SE-stained cells indicated by arrows in r-ctrl-injected skins, Fig. 3A, versus r-SCF-injected skins, Fig 3B). We confirmed SE- positive cells in rSCF-treated skins also possessed round-oval nuclei (Fig. 3C). Maximum mast cell enrichment was observed 12 hours post injection (hpi) of rSCF (Fig. 3D). When we challenged X. laevis with Bd, we did not detect differences in skin fungal loads between control (r-ctrl- injected) and mast cell-enriched groups at 7 days post infection (dpi; Fig. 3E). However, fungal loads were almost two-fold lower in mast cell- compared to control-enriched animals at 21 dpi, allowing for approximately four cycles of zoospore to mature zoosporangia development (Fig. 3E).

Enriching frog cutaneous mast cells lowers Bd loads.

Representative images of SE stained (A) control and (B) mast cell-enriched skin 12 hpi. (C) We confirmed the enriched population was composed of mono-morphonuclear cells. (D) Mast cell enrichment was optimized across several time points by quantifying SE-positive cells per field of view under 40x magnification. Results represent means ± SEM from 3 animals per time point (2 experimental repeats). (E) Mast cell-enriched and control dorsal skins were collected from X. laevis 7- and 21-dpi. Bd loads are represented as the number of zoospore genomic equivalents (GE) x 104 per μl of total input DNA. Time points were analyzed independently. Results represent means ± SEM from 7 animals per experimental group (N=7). Asterisks indicate significance: p < 0.05 by (D) one-way ANOVA with Tukey post-hoc analysis or (E) Student’s t-test.

Mast cells protect frogs from Bd-elicited inflammation-associated pathology

To explore potential mechanisms of mast cell-mediated protection against Bd, we compared the gene expression profiles of r-ctrl- and rSCF-administered Bd-infected frog skins at 21 dpi. Among the top differentially expressed genes, we noted mast cell-enriched, Bd-infected skins possessed greater transcripts for genes associated with cutaneous strength and integrity (lamc2), epidermal cell maturation, lymphocyte recruitment (b3gnt3.1)29, 30, 31, as well as ion and nutrient flow (gjb3l). (Fig. 4A). Moreover, mast cell-enriched Bd-challenged skin exhibited greater expression of genes associated with protection of the mucosa and epithelial healing (ttf3.6s)32, pathogen nutrient deprivation in mucus (slc1a5s)33, and mucus production (duoxa1.s34; Fig. 4A). In striking contrast, skins from control Bd-infected frogs revealed greater expression of genes associated with leukocyte infiltration and inflammation (e.g., ccl19, cxcl16, adamt13, csf3r; Fig. 4A). These transcriptional profiles were supported by our histological observations wherein control Bd- infected skins exhibited hyperkeratosis, epidermal hyperplasia, jagged stratum corneum, and extensive leukocyte infiltration (Fig. 4B), while mast cell-enriched Bd-infected tissues appeared largely devoid of these pathologies (Fig. 4C).

Consequences of cutaneous mast cell and neutrophil enrichment.

(A) RNAseq analysis of skin tissue from control (r-ctrl) or mast cell-enriched (rSCF) Bd-infected X. laevis at 21 dpi. Heat map of the top 30 DEGs, numbers matched to colors represent log2 fold change in expression. (B & C) Representative images of control and mast cell-enriched, Bd-infected skins, 21 dpi, demonstrating differences in epidermal thickening. (D & E) Representative images of control (r-ctrl) and neutrophil-enriched (rCSF3) skins collected 12 hpi demonstrating differences in epidermal thickening. (F) Kinetics of neutrophil enrichment following subcutaneous rCSF3 administration. Results are means + SEM of SE-positive cells per field of view from 4 animals per time point (N=4). (G) Bd loads from control and neutrophil-enriched skin tissue 7 dpi, (N=6). Asterisks indicate significance: p < 0.05 by (F) one-way ANOVA with Tukey post-hoc analysis or (G) Student’s t-test.

Cutaneous neutrophil enrichment results in increased Bd fungal loads

Neutrophils are one of the first leukocytes to infiltrate infected tissues, typically amplifying inflammation35. All vertebrate neutrophils depend on CSF3 for their differentiation and function36, and our previous work has demonstrated the gene encoding the CSF3 receptor (csf3r) is a marker of X. laevis neutrophils37, 38. Because csf3r expression was markedly elevated in control over mast cell-enriched skins of infected frogs (Fig. 4A), we next assessed the outcomes of enriching skin for neutrophils via subcutaneous rCSF3 administration. We confirmed neutrophil enrichment peaked 12 hp rCSF3 injection (Fig. 4F) and resulted in a thickened epidermis in comparison with r-ctrl injected skins of otherwise healthy animals (i.e., no Bd; Fig. 4D, E). When challenged with Bd, frogs with neutrophil-enriched skin possessed significantly greater Bd loads than control frogs (Fig. 4G). This suggests inflammation may be exacerbating Bd infections.

In vitro analyses of frog mast cells responses to Bd

Amphibians rely heavily on skin-produced antimicrobial peptides (AMPs) for antifungal protection39, and mast cells produce antimicrobial AMPs40. Thus, we examined whether mast cells could be sources of such AMPs during Bd infections. As anticipated, mast cells, but not neutrophils challenged with Bd, upregulated their expression of the AMP-encoding genes, pgla and magainin (Fig. S1A).

Mammalian mast cells are recognized as potent producers of the pleotropic anti- inflammatory cytokine, interleukin-4 (IL4)41. Notably, X. laevis mast cells challenged in vitro with Bd significantly upregulated their il4 gene expression (Fig. S1A), whereas almost no il4 expression was detected from either unstimulated or Bd-challenged neutrophils. To examine whether the frog IL4 might offer anti-Bd protection, we generated the X. laevis IL4 in recombinant form (rIL4) and confirmed subcutaneous injection of this moiety augmented expression of genes typically activated by the mammalian IL442 (cd36, metalloproteinase inhibitor 3-timp3, and monoamine oxidase A- maoa; Fig. S1B). However, subcutaneous administration of rIL4 in Bd-infected frogs did not alter their fungal loads (Fig. S1C), suggesting IL4 is probably not a major mechanism responsible for the observed mast cell-mediated anti-Bd protection.

Enrichment of frog skin mast cells alters mucus gland composition

Because mast cell-enriched frog skin had greater expression of genes associated with mucosal tissue integrity and mucus production (Fig. 4A), we next investigated whether the anti-Bd protection identified in mast cell-enriched skins could be due, at least in part, to differences in mucus production. Interestingly, while mucus glands in mock- and Bd-infected control skin sections contained normal levels of mucus, those from both mock- and Bd-infected frogs that had been enriched for cutaneous mast cells were significantly more filled with higher levels of mucin (Fig. 5A-C).

Cutaneous mast cells promote filled mucus glands.

Control (r-ctrl-injected) or mast cell-enriched (rSCF-injected) X. laevis were mock-infected or challenged with Bd for 21 days. Mucin content was examined in cutaneous mucus glands with Alcian Blue/PAS stain. Representative images of Bd-infected (A) control and (B) mast cell-enriched animals. Mucus glands are denoted by ‘m’, and epithelia are denoted by ‘Ep’. (C) ImageJ software was used to determine the proportion of positive mucin staining within each mucus gland. Results are mean + SEM percent positive mucin staining per mucus gland. Letters indicate group mean significance: p < 0.05 for treatment effects and interaction by two-way ANOVA. Skin tissue pieces from six individual frogs were used per treatment group (N=6). (D) Microbial phyla distribution across groups. Low abundance phyla (< 5% relative abundance are not shown). 10 dpi (E) community composition (Jaccard distances shown with 80% confidence ellipses) differed among all treatments. (F) Relative abundance of Bd-inhibitory bacteria and (G) bacterial richness were examined in control and mast cell enriched frogs, 10 days post Bd or mock challenge. Letters above bars in C, F and G indicate statistically different groups.

Enrichment of frog skin mast cells alters skin microbial composition

Amphibian skin-produced mucus may offer antimicrobial protection and serve as a selective substratum for commensal microbes, many of which are antifungal9. We found no significant differences in direct Bd-killing capacities of mucus isolated from mock- or Bd-challenged control or mast cell-enriched frogs (Fig. S2). By contrast, we observed substantial differences in skin microbiomes, including changes in bacterial composition and richness as well as relative abundances of Bd-inhibitory bacteria (Fig. 5D-G, Fig. S3). A total of 1645 bacterial amplicon sequence variants (ASVs) were identified from 20 bacterial phyla, seven of which were predominant (Fig. 5D). Of these, Verrucomicrobiota were only present on uninfected animals, whereas Acidobacteriota was only seen after 21 dpi on both control and mast cell-enriched, infected animals (Fig. 5D).

At 10 dpi, mast cell-enrichment resulted in a nominal shift in community composition compared to control frogs (Fig. 5E). Notably, while control, Bd-infected animals exhibited a drastic shift in community composition, mast cell-enriched animals possessed substantially less deviated community composition (Fig. 5E), suggesting that these cells are somehow counteracting the adverse effects of Bd on microbiome structure. These mast cell-mediated effects persisted to 21 dpi (Fig. S3A).

At 10 dpi, mast cell-enriched and mock-infected frogs possessed significantly greater abundance of Bd-inhibitory bacteria such as Chyrseobacterium sp., compared to control, mock infected animals (Fig. 5F)43. This suggests that mast cells may promote skin flora composition that is more antifungal. Control (non-enriched) Bd-infected frogs possessed significantly greater abundance of Bd-inhibitory bacteria than all other treatment groups (Fig. 5F). While mast cell- enriched, Bd-infected frogs had lower abundance of Bd-inhibitory bacteria than control infected frogs, they possessed higher abundance of inhibitory taxa than uninfected control animals (Fig. 5F). The Bd-inhibitory bacteria seen in greater abundance on mast cell-enriched, Bd-infected animals, included Roseateles sp., Flavobacterium sp., and Kaistia sp. We did not see significant differences in Bd-inhibitory bacteria across the treatment groups at 21 dpi (Fig. S3B).

Mast cell-enriched uninfected frogs exhibited increased frog skin bacterial richness at 10 dpi (Fig. 5G). While control Bd-infected animals exhibited significantly reduced skin microbial richness, mast cell-enriched Bd-infected frogs did not exhibit such a reduction in bacterial richness (Fig. 5G), supporting the idea that mast cells may be counteracting the adverse effects of Bd on skin microbiome composition.

Discussion

Amphibian extinction rates far outpace those of any other vertebrate class44. It is now well- established that chytrid fungi are major contributors to these declines, and strikingly, are considered the greatest infectious disease threat to biodivsersity45. The development of effective mitigation strategies though, is hindered by incomplete understanding of amphibian immune defenses and skin integrity. In this respect, while mast cells are recognized as key immune sentinels of tissues such as skin46, relatively little is known about this cell lineage outside of mammals. Our findings provide the first in-depth functional analyses of these cells in amphibians and explore their protective mechanisms during chytrid infections. This work presents a unique perspective on the evolution of mast cell functionality and will serve as a new avenue to explore ways to counteract the amphibian declines.

Mammalian mast cells directly interact with neurons innervating the skin47, and mucus production in mammals is tightly controlled by neuronal signaling48. In turn, mucus serves as a critical substratum for microbiota at mucosal barriers49. Here, we demonstrate that frog skin mast cells appear to confer significant protection against Bd by counteracting Bd-elicited inflammation, maintaining skin integrity, promoting mucus production, and preventing the disassembly of established skin microbiomes. Amphibian skin is much thinner and more permeable than that of mammals7, and as such, arguably represents a more penetrable barrier to pathogens. Because mammalian skin is relatively impermeable, mast cells are absent from mammalian epidermis and are instead found exclusively in their dermal layers50. In contrast, we have shown here and have observed across several classes of amphibians16 that mast cells are found in both epidermal and dermal layers. This localization has presumably evolved to support the more intimate contact between amphibian skin and their environments. Considering the importance of amphibian cutaneous integrity to their physiology51, it is likely that skin-resident mast cells co-evolved to support skin immunity, physiology, and symbiotic microbiota. Ongoing research continues to reveal functional differences between mammalian connective tissue and mucosa-resident mast cells52, and we suspect that there is similarly much to learn about the distinct physiological and immune roles of amphibian epidermal and dermal mast cells.

Bd infections caused major reductions in bacterial taxa richness, changes in composition and substantial increases in the relative abundance of Bd-inhibitory bacteria early in the infection. Similar changes to microbiome structure occur during experimental Bd infections of red-backed salamanders and mountain yellow-legged frogs8, 53. In turn, progressing Bd infections corresponded with a return to baseline levels of Bd-inhibitory bacteria abundance and rebounding microbial richness, albeit with dissimilar communities to those seen in control animals. These temporal changes indicate that amphibian microbiomes are dynamic, as are the effects of Bd infections on them. Indeed, Bd infections may have long-lasting impacts on amphibian microbiomes8. While Bd infections manifested in these considerable changes to frog skin microbiome structure, mast cell enrichment appeared to counteract these deleterious effects to their microbial composition. Presumably, the greater skin mucosal integrity and mucus production observed after mast cell enrichment served to stabilize the cutaneous environment during Bd infections, thereby ameliorating the Bd-mediated microbiome changes. While this work explored the changes in established antifungal flora, we anticipate the mast cell-mediated inhibition of Bd may be due to additional, yet unidentified bacterial or fungal taxa. Intriguingly, while mammalian skin mast cell functionality depends on microbiome elicited SCF production by keratinocytes54, our results indicate that frog skin mast cells in turn impact skin microbiome structure and likely their function. It will be interesting to further explore the interdependent nature of amphibian skin microbiomes and resident mast cells.

In contract to the mast cell effects, enrichment of neutrophils seemed to result in greater Bd skin burdens. This suggests that neutrophils and the cytokines they release may result in a greater level of inflammation, which is not protective. This is consistent with other studies that suggest that a strong immune response in the skin compartment of some species may be detrimental55, 56

It has become apparent that amphibian host-Bd interactions are highly complex and multifaceted while different amphibian species exhibit marked differences in their susceptibilities to this devastating pathogen2, 4. The findings described here emphasize the importance of skin- resident mast cells for successful anti-Bd defenses and demonstrate that these immune sentinels are intimately linked to many aspects of frog skin physiology. Our results indicate that when mast cells are enriched, the ensuing changes in the skin allow for greater resistance to infection by developing zoospores. Presumably, distinct amphibian species have evolved disparate interconnections between their skin mast cells and their cutaneous defenses, as dictated by their respective physiological and environmental pressures. In turn, these species-specific differences likely dictate whether and to what extent the skin-resident mast cells of a given amphibian species recognize and appropriately respond to Bd infections. We postulate that such differences may contribute to the differences in susceptibilities of amphibian species to pathogens like chytrid fungi.

Methods

Animals

Outbred 1 year-old (1.5-2”), mixed sex X. laevis were purchased from Xenopus 1 (Dexter, MI). All animals were housed and handled under strict laboratory regulations as per GWU IACUC (Approval number 15-024).

Recombinant cytokines and bone marrow granulocyte cultures

The X. laevis rSCF, rCSF3 and rIL4 were generated as previously described for rCSF357 and as detailed in the supplemental materials.

Bone marrow isolation, culture conditions, and establishment of neutrophil cultures have been previously described58. Mast cell cultures were generated according to protocols adapted from Koubourli et al. (2018) and Meurer et al. (2016) 59, 60. Isolated bone marrow cells were treated with 250 ng/µl of rSCF on Day 0, Day 4, Day 7, and collected for further analysis on Day 9. Cell cultures were maintained at 27°C with 5% CO2 in amphibian medium supplemented with 10% fetal bovine serum and 0.25% X. laevis serum. Neutrophil-like granulocytes were generated as above but with 250 ng/µl of rCSF3 on Day 0, Day 3, and collected for further analysis on Day 5. Cell cultures were maintained at 27°C with 5% CO2 in amphibian serum-free medium supplemented with 10% fetal bovine serum, 0.25% X. laevis serum, 10 μg/mL gentamicin (Thermo Fisher Scientific, Waltham, Massachusetts, USA), 100 U/mL penicillin, and 100 μg/mL streptomycin (Gibco, Thermo Fisher Scientific).

Enrichment of skin granulocyte subsets

Animals were subcutaneously injected between the skin and muscle layers with 5 µg/animal of rSCF, rCSF3, or r-ctrl in 5 µL of saline using finely pulled glass needles. Optimal time course and dose for in vivo mast cell and neutrophil enrichment were determined during preliminary experiments.

Recombinant interleukin-4 treatment

The capacity of the recombinant interleukin-4 (rIL4) to induce expression of genes associated with mammalian IL4 responses were assessed by injecting frogs (N=6 per treatment group subcutaneously with rIL4 (5 µg/animal) or r-ctrl in 5 µl of saline. After 6hrs, animals were sacrificed, and skins were isolated for gene expression analyses.

To examine the effect of rIL4 on Bd loads, frogs were infected with Bd by water bath (105 zoospores/mL) as described below and 1 day later injected subcutaneously, dorsally with rIL4 (5 µg/animal) or r-ctrl in 5 µl of saline. After an additional 9 days of infection, animals were sacrificed and their dorsal skin Bd loads examined.

Bd stocks and fungal challenge

Bd isolate JEL 197 was grown in 1% tryptone broth or on 1% tryptone agar plates (Difco Laboratories, Detroit, MI) supplemented with 100 U/mL penicillin and 100 μg/mL streptomycin (Gibco) at 19°C.

In vitro Bd killing was evaluated by incubating live Bd (maturing zoosporangia) with mast cells or neutrophils at ratios of 5:1 or 1:1 Bd cells per granulocyte. Cells were incubated at 27°C for three days before fungal loads were analyzed by absolute qPCR. Experimental groups were compared to pathogen DNA amounts derived from equal quantities of live Bd plated alone in otherwise identical conditions.

For in vivo infection studies, zoospores were harvested by flooding confluent tryptone agar plates with 2 mL sterile A-PBS for 10 minutes. Twelve hours post rSCF, rCSF3, rIL4, or r-ctrl injection, animals were infected with 107 zoospores or mock-infected in 100 mL of water (105 zoospores/mL). After 3 hrs, 400 mL of water was added to each tank. Skins were collected for histology and gene expression analysis on 1, 7, and 21 dpi.

Histology

X. laevis leukocyte cytology and cutaneous SE staining has been described57 and is detailed in the supplemental materials. An Alcian Blue/PAS staining kit (Newcomer Supply, Middleton, WI) was used to quantify mucin content from in vivo experiments. Histological analyses are detailed in the supplemental materials.

Electron Microscopy

Processing and imaging of cells for transmission and scanning electron microscopy (TEM and SEM) was conducted at the GWU Nanofabrication and Imaging Center (GWNIC). For transmission electron microscopy, cells were fixed as monolayers on six-well plates with 2.5% glutaraldehyde and 1% paraformaldehyde in 0.1 M sodium cacodylate buffer for one hour. Cells were treated with 1% osmium tetroxide in 0.1 M sodium cacodylate buffer for 1 hr. Following washes, cells were en bloc stained with 1% uranyl acetate in water overnight at 4°C. Samples were dehydrated through an ethanol series and embedded in epoxy resin using LX112. Inverted Beem capsules were placed into each tissue culture well to create on face blockfaces for sectioning. Resin was cured for 48 hrs at 60°C. The 95 nm sections were post-stained with 1% aqueous uranyl acetate and Reynold’s lead citrate. All imaging was performed at 80 KV in a Talos 200X transmission electron microscope (Thermo Fisher Scientific, Hillsboro, OR).

For SEM, cells were fixed with 2.5% glutaraldehyde / 1% paraformaldehyde in Sodium cacodylate buffer, followed by 1% OsO4, then dehydrated through an ethyl alcohol series. Coverslips were critical point dried and coated with 2 nm iridium. Cells were imaged using a Teneo Scanning Electron Microscope (Thermo Fisher Scientific).

Analyses of immune gene expression and Bd skin loads

These analyses have been described57 and are detailed in supplemental materials.

RNA sequencing

For transcriptomic profiling, bone marrow-derived neutrophil and mast cell cultures were generated as described above and FACS-sorted according to preestablished size and internal complexity parameters to isolate the respective subsets for further analyses. Sorted cells were immediately processed to extract and purify RNA. Flash frozen samples were sent to Azenta Life Sciences for all library preparation, RNA sequencing, and analyses. In short, polyadenylated RNA was isolated using Oligo dT beads. Enriched mRNAs were then fragmented for first and second strand cDNA synthesis. cDNA fragments were end repaired, 5’ phosphorylated, and dA-tailed.

Fragments were then ligated to universal adaptors and PCR-amplified. 150-bp paired-end sequencing was performed on an Illumina HiSeq platform.

FastQC was used to evaluate raw data quality. Adaptors sequences and poor-quality nucleotides were removed from reads using Trimmomatic v.0.36. The STAR aligner v.2.55.2b was used to map these reads to the Xenopus_laevis_9_2 reference genome from ENSEMBL. To determine differential gene expression, featureCount (Subread package v.1.5.2) was first used to count unique gene hits, which were then used with DESeq2 to calculate absolute log2-fold change.

Skin microbiome analyses

Towards microbiome studies, frogs were housed individually (N=6/treatment group). At indicated times, frogs were gently rinsed with sterile deionized water to remove transient microbes and gently swabbed 20 times, dorsally. Genomic DNA was extracted from swabs using a PowerSoil Pro kit on a Qiacube HT (QIAGEN, MD). One-step PCR library prep and dual-index paired-end Illumina sequencing was used to sequence the skin microbiome of individual frogs. A ∼380 base pair region in the V3-V5 region of the 16S rRNA gene using the universal primers 515F-Y (GTGYCAGCMGCCGCGGTAA) and 939R (CTTGTGCGGGCCCCCGTCAATTC) was used for amplification. Negative and positive controls (ZymoBIOMICS D6300 & D6305, Zymo, CA) were included in each round of extraction and PCR. Reactions were done in duplicate for each sample, pooled, cleaned with in-house Speed-beads (in a PEG/NaCl buffer), quantified with a Qubit4 (Invitrogen, MA) and pooled into a final library in equimolar proportion. The pooled library was sequenced on two Illumina MiSeq runs (v3 chemistry: 2x300 bp kit) at the Center for Conservation Genomics, Smithsonian National Zoo & Conservation Biology Institute.

Analyses were performed in the R environment version 4.0.3 (R Core Team, 2020) using methods detailed in the supplemental materials.

Statistical analyses

Differences in transcript expression were calculated with one-way or multi-way ANOVAs followed by Tukey post-hoc tests. Student’s t-tests were used to determine differences in Bd loads between treatments only. Statistical differences in mucin content and mucosome Bd-killing were assessed with the two-way ANOVA calculator available online through Statistics Kingdom. For RNA sequencing, p-values were calculated with the Wald test and were adjusted using the Benjamini-Hochberge procedure.

eLife Assessment

Overall, this is a significant study, and it is able to highlight mast cells in amphibians and their putative capability to respond to and combat fungal infections. Therefore, this study is important for the field. However, the manuscript is incomplete from the standpoint that there is functional data lacking on how these mast cells are activated and their precise functional properties. Such experiments would add substantial impact and rigor and fully support the conclusions.

Reviewer #1 (Public Review):

Summary:

The global decline of amphibians is primarily attributed to deadly disease outbreaks caused by the chytrid fungus, Batrachochytrium dendrobatidis (Bd). It is unclear whether and how skin-resident immune cells defend against Bd. Although it is well known that mammalian mast cells are crucial immune sentinels in the skin and play a pivotal role in the immune recognition of pathogens and orchestrating subsequent immune responses, the roles of amphibian mast cells during Bd infections are largely unknown. The current study developed a novel way to enrich X. laevis skin mast cells by injecting the skin with recombinant stem cell factor (SCF), a KIT ligand required for mast cell differentiation and survival. The investigators found an enrichment of skin mast cells provides X. laevis substantial protection against Bd and mitigates the inflammation-related skin damage resulting from Bd infection. Additionally, the augmentation of mast cells leads to increased mucin content within cutaneous mucus glands and shields frogs from the alterations to their skin microbiomes caused by Bd.

Strengths:

This study underscores the significance of amphibian skin-resident immune cells in defenses against Bd and introduces a novel approach to examining interactions between amphibian hosts and fungal pathogens.

Weaknesses:

The main weakness of the study is the lack of functional analysis of X. laevis mast cells. Upon activation, mast cells have the characteristic feature of degranulation to release histamine, serotonin, proteases, cytokines, and chemokines, etc. The study should determine whether X. laevis mast cells can be degranulated by two commonly used mast cell activators IgE and compound 48/80 for IgE-dependent and independent pathways. This can be easily done in vitro. It is also important to assess whether in vivo these mast cells are degranulated upon Bd infection using avidin staining to visualize vesicle releases from mast cells. Figure 3 only showed rSCF injection caused an increase in mast cells in naïve skin. They need to present whether Bd infection can induce mast cell increase and rSCF injection under Bd infection causes a mast cell increase in the skin. In addition, it is unclear how the enrichment of mast cells provides protection against Bd infection and alternations to skin microbiomes after infection. It is important to determine whether skin mast cells release any contents mentioned above.

Reviewer #2 (Public Review):

Summary:

In this study, Hauser et al investigate the role of amphibian (Xenopus laevis) mast cells in cutaneous immune responses to the ecologically important pathogen Batrachochytrium dendrobatidis (Bd) using novel methods of in vitro differentiation of bone marrow-derived mast cells and in vivo expansion of skin mast cell populations. They find that bone marrow-derived myeloid precursors cultured in the presence of recombinant X. laevis Stem Cell Factor (rSCF) differentiate into cells that display hallmark characteristics of mast cells. They inject their novel (r)SCF reagent into the skin of X. laevis and find that this stimulates the expansion of cutaneous mast cell populations in vivo. They then apply this model of cutaneous mast cell expansion in the setting of Bd infection and find that mast cell expansion attenuates the skin burden of Bd zoospores and pathologic features including epithelial thickness and improves protective mucus production and transcriptional markers of barrier function. Utilizing their prior expertise with expanding neutrophil populations in X. laevis, the authors compare mast cell expansion using (r)SCF to neutrophil expansion using recombinant colony-stimulating factor 3 (rCSF3) and find that neutrophil expansion in Bd infection leads to greater burden of zoospores and worse skin pathology.

Strengths:

The authors report a novel method of expanding amphibian mast cells utilizing their custom-made rSCF reagent. They rigorously characterize expanded mast cells in vitro and in vivo using histologic, morphologic, transcriptional, and functional assays. This establishes solid footing with which to then study the role of rSCF-stimulated mast cell expansion in the Bd infection model. This appears to be the first demonstration of the exogenous use of rSCF in amphibians to expand mast cell populations and may set a foundation for future mechanistic studies of mast cells in the X. laevis model organism.

Weaknesses:

The conclusions regarding the role of mast cell expansion in controlling Bd infection would be stronger with a more rigorous evaluation of the model, as there are some key gaps and remaining questions regarding the data. For example:

1. Granulocyte expansion is carefully quantified in the initial time courses of rSCF and rCSF3 injections, but similar quantification is not provided in the disease models (Figures 3E, 4G, 5D-G). A key implication of the opposing effects of mast cell vs neutrophil expansion is that mast cells may suppress neutrophil recruitment or function. Alternatively, mast cells also express notable levels of csfr3 (Figure 2) and previous work from this group (Hauser et al, Facets 2020) showed rG-CSF-stimulated peritoneal granulocytes express mast cell markers including kit and tpsab1, raising the question of what effect rCSF3 might have on mast cell populations in the skin. Considering these points, it would be helpful if both mast cells and neutrophils were quantified histologically (based on Figure 1, they can be readily distinguished by SE or Giemsa stain) in the Bd infection models.

2. Epithelial thickness and inflammation in Bd infection are reported to be reduced by rSCF treatment (Figure 3E, 5A-B) or increased by rCSF3 treatment (Figure 4G) but quantification of these critical readouts is not shown.

3. Critical time points in the Bd model are incompletely characterized. Mast cell expansion decreases zoospore burden at 21 dpi, while there is no difference at 7 dpi (Figure 3E). Conversely, neutrophil expansion increases zoospore burden at 7 dpi, but no corresponding 21 dpi data is shown for comparison (Figure 4G). Microbiota analysis is performed at a third time point,10 dpi (Figure 5D-G), making it difficult to compare with the data from the 7 dpi and 21 dpi time points. Reporting consistent readouts at these three time points is important to draw solid conclusions about the relationship of mast cell expansion to Bd infection and shifts in microbiota.

4. Although the effect of rSCF treatment on Bd zoospores is significant at 21 dpi (Figure 3E), bacterial microbiota changes at 21 dpi are not (Figure S3B-C). This discrepancy, how it relates to the bacterial microbiota changes at 10 dpi, and why 7, 10, and 21 dpi time points were chosen for these different readouts (Figure 5F-G), is not discussed.

5. The time course of rSCF or rCSF3 treatments relative to Bd infection in the experiments is not clear. Were the treatments given 12 hours prior to the final analysis point to maximize the effect? For example, in Figure 3E, were rSCF injections given at 6.5 dpi and 20.5 dpi? Or were treatments administered on day 0 of the infection model? If the latter, how do the authors explain the effects at 7 dpi or 21 dpi given mast cell and neutrophil numbers return to baseline within 24 hours after rSCF or rCSF3 treatment, respectively?

The title of the manuscript may be mildly overstated. Although Bd infection can indeed be deadly, mortality was not a readout in this study, and it is not clear from the data reported that expanding skin mast cells would ultimately prevent progression to death in Bd infections.

Reviewer #3 (Public Review):

Summary:

Hauser et al. provide an exceptional study describing the role of resident mast cells in amphibian epidermis that produce anti-inflammatory cytokines that prevent Batrachochytrium dendrobatidis (Bd) infection from causing harmful inflammation, and also protect frogs from changes in skin microbiomes and loss of mucin in glands and loss of mucus integrity that otherwise cause changes to their skin microbiomes. Neutrophils, in contrast, were not protective against Bd infection. Beyond the beautiful cytology and transcriptional profiling, the authors utilized elegant cell enrichment experiments to enrich mast cells by recombinant stem cell factor, or to enrich neutrophils by recombinant colony-stimulating factor-3, and examined respective infection outcomes in Xenopus.

Strengths:

Through the use of recombinant IL4, the authors were able to test and eliminate the hypothesis that mast cell production of IL4 was the mechanism of host protection from Bd infection. Instead, impacts on the mucus glands and interaction with the skin microbiome are implicated as the protective mechanism. These results will press disease ecologists to examine the relative importance of this immune defense among species, the influence of mast cells on the skin microbiome and mucosal function, and open the potential for modulating mucosal defense.

Weaknesses:

A reduction of bacterial diversity upon infection, as described at the end of the results section, may not always be an "adverse effect," particularly given that anti-Bd function of the microbiome increased. Some authors (see Letourneau et al. 2022 ISME, or Woodhams et al. 2023 DCI) consider these short-term alterations as encoding ecological memory, such that continued exposure to a pathogen would encounter an enriched microbial defense. Regardless, mast cell-initiated protection of the mucus layer may negate the need for this microbial memory defense.

While the description of the mast cell location in the epidermal skin layer in amphibians is novel, it is not known how representative these results are across species ranging in chytridiomycosis susceptibility. No management applications are provided such as methods to increase this defense without the use of recombinant stem cell factor, and more discussion is needed on how the mast cell component (abundance, distribution in the skin) of the epidermis develops or is regulated.

None

Author response:

The following is the authors’ response to the current reviews.

Reviewer #1 (Public Review):

Summary: The global decline of amphibians is primarily attributed to deadly disease outbreaks caused by the chytrid fungus, Batrachochytrium dendrobatidis (Bd). It is unclear whether and how skin-resident immune cells defend against Bd. Although it is well known that mammalian mast cells are crucial immune sentinels in the skin and play a pivotal role in immune recognition of pathogens and orchestrating subsequent immune responses, the roles of amphibian mast cells during Bd infections is largely unknown. The current study developed a novel way to enrich X. laevis skin mast cells by injecting the skin with recombinant stem cell factor (SCF), a KIT ligand required for mast cell differentiation and survival. The investigators found an enrichment of skin mast cells provides X. laevis substantial protection against Bd and mitigates the inflammation-related skin damage resulting from Bd infection. Additionally, the augmentation of mast cells leads to increased mucin content within cutaneous mucus glands and shields frogs from the alterations to their skin microbiomes caused by Bd.

Strengths: This study underscores the significance of amphibian skin-resident immune cells in defenses against Bd and introduces a novel approach to examining interactions between amphibian hosts and fungal pathogens.

We thank the reviewer for recognizing the significance and the novelty of our work.

Weaknesses: The main weakness of the study is lack of functional analysis of X. laevis mast cells. Upon activation, mast cells have the characteristic feature of degranulation to release histamine, serotonin, proteases, cytokines, and chemokines, etc. The study should determine whether X. laevis mast cells can be degranulated by two commonly used mast cell activators IgE and compound 48/80 for IgE-dependent and independent pathway. This can be easily done in vitro. It is also important to assess whether in vivo these mast cells are degranulated upon Bd infection using avidin staining to visualize vesicle releases from mast cells. Figure 3 only showed rSCF injection caused an increase in mast cells in naïve skin. They need to present whether Bd infection can induce mast cell increase and rSCF injection under Bd infection causes a mast cell increase in the skin. In addition, it is unclear how the enrichment of mast cells provides the protection against Bd infection and alternations to skin microbiomes after infection. It is important to determine whether skin mast cell release any contents mentioned above.

We would like to thank the reviewer for taking the time to review our work and providing us with valuable feedback.

Please note, that as indicated in our previous rebuttal to reviewers, amphibians do not possess the IgE antibody isotype1.

To our knowledge, there are no published works describing the approaches used in studying mammalian mast cell degranulation towards examining amphibian mast cells. While there are commercially available kits and reagents for examining mammalian mast cell granule content, most of these do not cross-react with amphibian counterparts. This is especially true of cytokines and chemokines, which diverged quickly with evolution and thus do not share substantial protein sequence identity across species as diverged as frogs and mammals. We would also like to highlight the fact that several studies suggest that amphibian mast cells lack histamine2, 3, 4, 5 and serotonin2, 6. While following up on these findings would be possible, we would like to respectfully emphasize that adopting approaches used in mammalian research to comparative immunology work is not always straightforward.

As we highlight in our manuscript, frog mast cells upregulate their expression of interleukin-4 (IL4), a hallmark cytokine associated with mammalian mast cells7. The additional findings presented in our revised manuscript indicate that mast cells respond to Bd by upregulating IL4 expression in vitro and in vivo. Together, this suggests that IL4 may be a central means by which frog mast cells confer protection against Bd, by counteracting Bd-elicited inflammation, including minimizing neutrophil infiltration, maintaining skin integrity, and promoting cutaneous mucus production. Please find that these additional results are presented in Figure 8 and are described in the results and discussion sections of our revised manuscript.

Our attempts to elicit degranulation of frog mast cells using compound 48/80 have so far not been successful. This may reflect technical issues with assays optimized for mammalian mast cells or biological difference between frog and mammalian mast cells, such as species differences in mas-related G-protein coupled receptors, through which compound 48/80 acts8. We will continue to explore means to study frog mast cell degranulation both in vitro and in vivo but also respectfully point out that while degranulation is a feature commonly associated with mammalian mast cells, this is not the only means by which the mammalian mast cells confer their immunological effects. Indeed, our studies suggest that frog mast cell IL4 production may be a key means by which these cells offer anti-Bd protection.

Please note that we successfully adopted an avidin staining approach to visualize mast cell heparin content in vitro and to evaluate cutaneous mast cell numbers in vivo in control and mast cell-enriched, mock- and Bd-infected animals. This additional work is depicted in Figure 4 and addressed in the results and discussion sections of our revised manuscript.

Reviewer #2 (Public Review):

Summary: In this study, Hauser et al investigate the role of amphibian (Xenopus laevis) mast cells in cutaneous immune responses to the ecologically important pathogen Batrachochytrium dendrobatidis (Bd) using novel methods of in vitro differentiation of bone marrow-derived mast cells and in vivo expansion of skin mast cell populations. They find that bone marrow-derived myeloid precursors cultured in the presence of recombinant X. laevis Stem Cell Factor (rSCF) differentiate into cells that display hallmark characteristics of mast cells. They inject their novel (r)SCF reagent in the skin of X. laevis and find that this stimulates expansion of cutaneous mast cell populations in vivo. They then apply this model of cutaneous mast cell expansion in the setting of Bd infection and find that mast cell expansion attenuates skin burden of Bd zoospores and pathologic features including epithelial thickness and improves protective mucus production and transcriptional markers of barrier function. Utilizing their prior expertise with expanding neutrophil populations in X. laevis, the authors compare mast cell expansion using (r)SCF to neutrophil expansion using recombinant colony stimulating factor 3 (rCSF3) and find that neutrophil expansion in Bd infection leads to greater burden of zoospores and worse skin pathology. Combining these two observations, they demonstrate that mast cell expansion using rSCF attenuates cutaneous neutrophilic infiltration. They further show that mast cell expansion correlates to cutaneous IL-4 expression, and that treatment with exogenous rIL-4 reduces neutrophilic infiltration and restores markers of epithelial health, offering a mechanism by which mast cell expansion protects from Bd infection.

Strengths: The authors report a novel method of expanding amphibian mast cells utilizing their custom-made rSCF reagent. They rigorously characterize expanded mast cells in vitro and in vivo using histologic, morphologic, transcriptional, and functional assays. This establishes solid footing with which to then study the role of rSCF-stimulated mast cell expansion in the Bd infection model. This appears to be the first demonstration of exogenous use of rSCF in amphibians to expand mast cell populations and may set a foundation for future mechanistic studies of mast cells in the X. laevis model organism. Building on prior work, they are able to contrast mast cell expansion with their neutrophil expansion model, allowing them to infer a mechanistic link between mast cell expansion and IL-4 production and subsequent suppression of neutrophil infiltration and cutaneous dysbiosis.

We thank the reviewer for recognizing the rigorousness and utility of the studies presented in our manuscript.

Weaknesses: The main weaknesses derive from technical limitations inherent to the Xenopus model at this time. For example, in mice a mechanistic study would be expected to use IL-4 knockouts, preferably mast cell-specific, to prove the link between mast cell expansion and IL-4 production being necessary and sufficient to suppress neutrophils. However, the novel reagents in this manuscript present a compelling technical advance and a step forward in the tools available to study amphibian biology.

We agree with the reviewer that an IL4 knock-out animal model would be a great way to support our findings. Unfortunately, working with a non-mammalian model such as X. laevis poses limitations that include lack of knock-out lines for immunology research. Moreover, as mentioned in our manuscript, we do not believe that IL4 is the sole mast cell-produced component responsible for the conferred antifungal protection. We thank the reviewer for acknowledging the limitations of our model system and recognizing the novelty, technical advances, and merits of the work presented in our manuscript.

In addition to their discussion, one open question from the revised manuscript is how a single treatment with rSCF leads to a peak in mast cell numbers and then decline to baseline in mock-infected frogs, while Bd infection either sustains rSCF-boosted mast cells or leads to steady mast cell increase over time in control-treated frogs. Whether this is mediated by endogenous SCF or some other factor remains unexplored.

This is an interesting question that we hope to explore in future studies. We did not see significant differences in skin SCF gene expression at 21 days post Bd infection. This does not rule out the possibility that the observed Bd-mediated effects to frog skin mast cell composition are not due to changes in skin SCF gene expression at earlier infection times, alone or in combination with other host or pathogen derived factors. We know that other factors are responsible for homing/retention of antimicrobial and immunosuppressive granulocyte subsets within frog skin9 and we postulate that some of these may be distinct mast cell types. Additionally, Bd is known to produce a myriad of immunomodulatory factors10, which may well also directly affect frog skin mast cell composition. Mammalian mast cells are heterogenous and are homed or recruited into tissues by an extensive array of host as well as microbiome-derived components11, 12. Undoubtedly, the frog skin mast cell composition is likewise complex, dynamic, and contingent on a plethora of host, cutaneous microbial flora- and in this case also Bd-produced factors.

References

(1) Flajnik, M.F. A cold-blooded view of adaptive immunity. Nat Rev Immunol 18, 438-453 (2018).

(2) Mulero, I., Sepulcre, M.P., Meseguer, J., Garcia-Ayala, A. & Mulero, V. Histamine is stored in mast cells of most evolutionarily advanced fish and regulates the fish inflammatory response. Proc Natl Acad Sci U S A 104, 19434-19439 (2007).

(3) Reite, O.B. A phylogenetical approach to the functional significance of tissue mast cell histamine. Nature 206, 1334-1336 (1965).

(4) Reite, O.B. Comparative physiology of histamine. Physiol Rev 52, 778-819 (1972).

(5) Takaya, K., Fujita, T. & Endo, K. Mast cells free of histamine in Rana catasbiana. Nature 215, 776-777 (1967).

(6) Galli, S.J. New insights into "the riddle of the mast cells": microenvironmental regulation of mast cell development and phenotypic heterogeneity. Lab Invest 62, 5-33 (1990).

(7) Babina, M., Guhl, S., Artuc, M. & Zuberbier, T. IL-4 and human skin mast cells revisited: reinforcement of a pro-allergic phenotype upon prolonged exposure. Archives of dermatological research 308, 665-670 (2016).

(8) Hermans, M.A.W. et al. Human Mast Cell Line HMC1 Expresses Functional Mas-Related G-Protein Coupled Receptor 2. Front Immunol 12, 625284 (2021).

(9) Hauser, K. et al. Discovery of granulocyte-lineage cells in the skin of the amphibian Xenopus laevis. FACETS 5, 571 (2020).

(10) Rollins-Smith, L.A. & Le Sage, E.H. Batrachochytrium fungi: stealth invaders in amphibian skin. Curr Opin Microbiol 61, 124-132 (2021).

(11) Halova, I., Draberova, L. & Draber, P. Mast cell chemotaxis - chemoattractants and signaling pathways. Front Immunol 3, 119 (2012).

(12) West, P.W. & Bulfone-Paus, S. Mast cell tissue heterogeneity and specificity of immune cell recruitment. Front Immunol 13, 932090 (2022).

The following is the authors’ response to the original reviews.

Reviewer #1 (Public Review):

Summary:

The global decline of amphibians is primarily attributed to deadly disease outbreaks caused by the chytrid fungus, Batrachochytrium dendrobatidis (Bd). It is unclear whether and how skin-resident immune cells defend against Bd. Although it is well known that mammalian mast cells are crucial immune sentinels in the skin and play a pivotal role in the immune recognition of pathogens and orchestrating subsequent immune responses, the roles of amphibian mast cells during Bd infections are largely unknown. The current study developed a novel way to enrich X. laevis skin mast cells by injecting the skin with recombinant stem cell factor (SCF), a KIT ligand required for mast cell differentiation and survival. The investigators found an enrichment of skin mast cells provides X. laevis substantial protection against Bd and mitigates the inflammation-related skin damage resulting from Bd infection. Additionally, the augmentation of mast cells leads to increased mucin content within cutaneous mucus glands and shields frogs from the alterations to their skin microbiomes caused by Bd.

Strengths:

This study underscores the significance of amphibian skin-resident immune cells in defenses against Bd and introduces a novel approach to examining interactions between amphibian hosts and fungal pathogens.

We thank the reviewer for acknowledging the novelty and importance of the work presented in our manuscript.

Weaknesses:

The main weakness of the study is the lack of functional analysis of X. laevis mast cells. Upon activation, mast cells have the characteristic feature of degranulation to release histamine, serotonin, proteases, cytokines, and chemokines, etc. The study should determine whether X. laevis mast cells can be degranulated by two commonly used mast cell activators IgE and compound 48/80 for IgE-dependent and independent pathways. This can be easily done in vitro. It is also important to assess whether in vivo these mast cells are degranulated upon Bd infection using avidin staining to visualize vesicle releases from mast cells. Figure 3 only showed rSCF injection caused an increase in mast cells in naïve skin. They need to present whether Bd infection can induce mast cell increase and rSCF injection under Bd infection causes a mast cell increase in the skin. In addition, it is unclear how the enrichment of mast cells provides protection against Bd infection and alternations to skin microbiomes after infection. It is important to determine whether skin mast cells release any contents mentioned above.

We would like to thank the reviewer for taking the time to review our work and providing us with valuable feedback. We feel that we have successfully incorporated the reviewer’s suggestions into our revised manuscript, thereby improving this work.

Please note that amphibians do not possess the IgE antibody isotype1.

To our knowledge there have been no published work assimilating approaches used when studying mammalian mast cell degranulation towards examining amphibian mast cells. While there are commercially available kits and reagents for examining mammalian mast cell granule content, most of these reagents do not cross-react with amphibian counterparts. This is especially true of cytokines and chemokines, which diverged quickly with evolution and thus do not share substantial protein sequence identity across species as diverged as frogs and mammals. Additionally, several studies suggest that amphibian mast cells lack histamine2, 3, 4, 5 and serotonin2, 6. Respectfully, while following up on these findings is possible, we would not consider adopting approaches used in mammalian research to comparative immunology work as easy.

As noted in our manuscript, frog mast cells upregulate their expression of interleukin-4 (IL4), which is a hallmark cytokine associated with mammalian mast cells7. The additional findings, presented in our revised manuscript indicate that mast cells respond to Bd by upregulating IL4 expression in vitro and in vivo. In turn, our work indicates that IL4 may be a central means by which frog mast cells confer protection against Bd, by counteracting Bd-elicited inflammation, including minimizing neutrophil infiltration, maintaining skin integrity, and promoting mucus production by skin mucus glands. Please find that these additional findings are presented in Figure 8 of our revised manuscript and are described in the results and discussion sections of the paper.

Our attempts to elicit degranulation of frog mast cells using compound 48/80 have so far not been successful. This may reflect technical issues with assays optimized for mammalian mast cells or biological difference between frog and mammalian mast cells, such as species differences in mas-related G-protein coupled receptors, through which compound 48/80 acts8. We will continue explore means to study frog mast cell degranulation both in vitro and in vivo but would also like to respectfully point out that while mast cell degranulation is a feature most associated with mammalian mast cells, this is not the only means by which the mammalian mast cells confer their immunological effects. Indeed, our additional studies suggest that mast cell IL4 production may be a key means by which these cells offer anti-Bd protection.

Please find that we have adopted an avidin-staining approach to visualize mast cell heparin content in vitro and to evaluate mast cell numbers in vivo in the skins of control and mast cell-enriched, mock- and Bd-infected animals. This additional work is depicted in Figure 4 of our revised manuscript and addressed in the results and discussion sections of our revised paper.

Reviewer #2 (Public Review):

Summary:

In this study, Hauser et al investigate the role of amphibian (Xenopus laevis) mast cells in cutaneous immune responses to the ecologically important pathogen Batrachochytrium dendrobatidis (Bd) using novel methods of in vitro differentiation of bone marrow-derived mast cells and in vivo expansion of skin mast cell populations. They find that bone marrow-derived myeloid precursors cultured in the presence of recombinant X. laevis Stem Cell Factor (rSCF) differentiate into cells that display hallmark characteristics of mast cells. They inject their novel (r)SCF reagent into the skin of X. laevis and find that this stimulates the expansion of cutaneous mast cell populations in vivo. They then apply this model of cutaneous mast cell expansion in the setting of Bd infection and find that mast cell expansion attenuates the skin burden of Bd zoospores and pathologic features including epithelial thickness and improves protective mucus production and transcriptional markers of barrier function. Utilizing their prior expertise with expanding neutrophil populations in X. laevis, the authors compare mast cell expansion using (r)SCF to neutrophil expansion using recombinant colony-stimulating factor 3 (rCSF3) and find that neutrophil expansion in Bd infection leads to greater burden of zoospores and worse skin pathology.

Strengths:

The authors report a novel method of expanding amphibian mast cells utilizing their custom-made rSCF reagent. They rigorously characterize expanded mast cells in vitro and in vivo using histologic, morphologic, transcriptional, and functional assays. This establishes solid footing with which to then study the role of rSCF-stimulated mast cell expansion in the Bd infection model. This appears to be the first demonstration of the exogenous use of rSCF in amphibians to expand mast cell populations and may set a foundation for future mechanistic studies of mast cells in the X. laevis model organism.

We thank the reviewer for recognizing the breadth and extent of the undertaking that culminated in this manuscript. Indeed, this manuscript would not have been possible without considerable reagent development and adaptation of techniques that had previously not been used for amphibian immunity research. In line with the reviewer’s sentiment, to our knowledge this is the first report of using molecular approaches to augment amphibian mast cells, which we hope will pave the way for new areas of research within the fields of comparative immunology and amphibian disease biology.

Weaknesses:

The conclusions regarding the role of mast cell expansion in controlling Bd infection would be stronger with a more rigorous evaluation of the model, as there are some key gaps and remaining questions regarding the data. For example:

(1) Granulocyte expansion is carefully quantified in the initial time courses of rSCF and rCSF3 injections, but similar quantification is not provided in the disease models (Figures 3E, 4G, 5D-G). A key implication of the opposing effects of mast cell vs neutrophil expansion is that mast cells may suppress neutrophil recruitment or function. Alternatively, mast cells also express notable levels of csfr3 (Figure 2) and previous work from this group (Hauser et al, Facets 2020) showed rG-CSF-stimulated peritoneal granulocytes express mast cell markers including kit and tpsab1, raising the question of what effect rCSF3 might have on mast cell populations in the skin. Considering these points, it would be helpful if both mast cells and neutrophils were quantified histologically (based on Figure 1, they can be readily distinguished by SE or Giemsa stain) in the Bd infection models.

We thank the reviewer for this insightful suggestion. Please find that we successfully adopted an in situ hybridization approach to evaluate neutrophil numbers in the skins of control and mast cell-enriched, mock- and Bd-infected animals based on expression of the neutrophil marker, myeloperoxidase (MPO9). Please find these results are presented in Figures 6 and 8 of our revised manuscript and addressed in the appropriate sections of our revised paper.

Our findings suggest that rSCF administration results in the accumulation of mast cells that are polarized such, that they ablate the inflammatory response elicited by Bd infection, such as through mechanisms like IL4 production. Mammalian mast cells, including peritonea-resident mast cells, express csf3r10, 11. For this reason, we used MPO expression to visualize neutrophil skin infiltration in Figures 6 and 8 of our revised work. While the X. laevis animal model does not permit nearly the degree of immune cell resolution afforded by mammalian animal models, we do know that the adult X. laevis peritonea contain a myriad of immune cell subsets. We anticipate that the high kit expression reported by Hauser et al., 2020 in the rCSF3-recruited peritoneal leukocytes reflects the presence of mast cells therein.

Please find that we have used avidin-staining and MPO in situ hybridization to respectively visualize and enumerate mast cells and neutrophils in the skin of control and mast cell-enriched, mock- and Bd-infected animals. Indeed, our results show interesting, experimental condition-dependent changes in both the skin neutrophil and mast cell numbers. The results of these additional studies are presented in Figures 4, 6 and 8 of the revised manuscript and addressed in the results and discussions sections of our revised paper.

(2) Epithelial thickness and inflammation in Bd infection are reported to be reduced by rSCF treatment (Figure 3E, 5A-B) or increased by rCSF3 treatment (Figure 4G) but quantification of these critical readouts is not shown.

We thank the reviewer for this suggestion. We scored epithelial thickness under the distinct conditions described in our manuscript and presented the quantified data in Figures 5 and 8 of the revised paper.

(3) Critical time points in the Bd model are incompletely characterized. Mast cell expansion decreases zoospore burden at 21 dpi, while there is no difference at 7 dpi (Figure 3E). Conversely, neutrophil expansion increases zoospore burden at 7 dpi, but no corresponding 21 dpi data is shown for comparison (Figure 4G). Microbiota analysis is performed at a third time point,10 dpi (Figure 5D-G), making it difficult to compare with the data from the 7 dpi and 21 dpi time points. Reporting consistent readouts at these three time points is important to draw solid conclusions about the relationship of mast cell expansion to Bd infection and shifts in microbiota.

We thank reviewer for noting this discrepancy. Please find that we have repeated our mast cell-enrichment, Bd-challenge studies, examining days 10 and 21 post infection. Our new findings indicate that compared to control animals, mast cell-enrichment does result in significant reduction in Bd loads at both 10 and 21 dpi. The difference in Bd loads between r-ctrl and rSCF-treated animals at 10 dpi corroborates the other parameters that are altered between the two treatment groups at this experimental time point.

Our question regarding the roles of inflammatory granulocytes/neutrophils during Bd infections was that of ‘how’ rather ‘when’ these cells affect Bd infections. Thus, and because the central focus of this work was mast cells and not other granulocyte subsets; when we saw that rCSF3-recruited granulocytes adversely affect Bd infections at 7 days, we did not pursue the kinetics of these observations further. We plan to explore the roles of inflammatory mediators and immune cell subsets during the course of Bd infections but feel that these future studies are more peripheral to the central thesis of the present manuscript regarding the roles of frog mast cells during Bd infections.

(4) Although the effect of rSCF treatment on Bd zoospores is significant at 21 dpi (Figure 3E), bacterial microbiota changes at 21 dpi are not (Figure S3B-C). This discrepancy, how it relates to the bacterial microbiota changes at 10 dpi, and why 7, 10, and 21 dpi time points were chosen for these different readouts (Figure 5F-G), is not discussed.

Please find that our additional studies indicate that compared to control animals, frog skin mast cell-enrichment results in significant reduction in Bd loads at 10 dpi. This corroborate our other findings including the observation that at 10 dpi, control animals exhibit reduced microbial richness whereas mast cell-enriched frogs were protected from this disruption of their microbiome. The amphibian microbiome serves as a major barrier to these fungal infections12 and we anticipate that Bd-mediated disruption of microbial richness facilitates host skin colonization by this pathogen. In turn, we anticipate that frog mast cells are conferring the observed anti-Bd protection in part by preventing microbial disassembly and thus interfering with optimal Bd colonization and growth on frog skins. Please find that we acknowledge and discuss these notions in our revised manuscript.

(5) The time course of rSCF or rCSF3 treatments relative to Bd infection in the experiments is not clear. Were the treatments given 12 hours prior to the final analysis point to maximize the effect? For example, in Figure 3E, were rSCF injections given at 6.5 dpi and 20.5 dpi? Or were treatments administered on day 0 of the infection model? If the latter, how do the authors explain the effects at 7 dpi or 21 dpi given mast cell and neutrophil numbers return to baseline within 24 hours after rSCF or rCSF3 treatment, respectively?

Please find that in our revised manuscript, we underlined the kinetics of our animal treatments and Bd-infections. In brief, for mast cell-enrichment, animals were injected with r-ctrl or rSCF, challenged 12 hours later with Bd and examined after 10 (per reviewers’ suggestions) and 21 days of infection. For neutrophil enrichment, animals were injected with r-ctrl or rCSF3, challenged 12 hours later with Bd and examined after 7 days of infection.

The title of the manuscript may be mildly overstated. Although Bd infection can indeed be deadly, mortality was not a readout in this study, and it is not clear from the data reported that expanding skin mast cells would ultimately prevent progression to death in Bd infections.

We acknowledge this point. The revised manuscript will be titled: “Amphibian mast cells: barriers to chytrid fungus infections”.

Reviewer #3 (Public Review):

Summary:

Hauser et al. provide an exceptional study describing the role of resident mast cells in amphibian epidermis that produce anti-inflammatory cytokines that prevent Batrachochytrium dendrobatidis (Bd) infection from causing harmful inflammation, and also protect frogs from changes in skin microbiomes and loss of mucin in glands and loss of mucus integrity that otherwise cause changes to their skin microbiomes. Neutrophils, in contrast, were not protective against Bd infection. Beyond the beautiful cytology and transcriptional profiling, the authors utilized elegant cell enrichment experiments to enrich mast cells by recombinant stem cell factor, or to enrich neutrophils by recombinant colony-stimulating factor-3, and examined respective infection outcomes in Xenopus.

Strengths:

Through the use of recombinant IL4, the authors were able to test and eliminate the hypothesis that mast cell production of IL4 was the mechanism of host protection from Bd infection. Instead, impacts on the mucus glands and interaction with the skin microbiome are implicated as the protective mechanism. These results will press disease ecologists to examine the relative importance of this immune defense among species, the influence of mast cells on the skin microbiome and mucosal function, and open the potential for modulating mucosal defense.

We thank the reviewer for recognizing the utility of the work presented in our manuscript.

Weaknesses:

A reduction of bacterial diversity upon infection, as described at the end of the results section, may not always be an "adverse effect," particularly given that anti-Bd function of the microbiome increased. Some authors (see Letourneau et al. 2022 ISME, or Woodhams et al. 2023 DCI) consider these short-term alterations as encoding ecological memory, such that continued exposure to a pathogen would encounter an enriched microbial defense. Regardless, mast cell-initiated protection of the mucus layer may negate the need for this microbial memory defense.

We thank the reviewer their insightful comment. We have revised our discussion to include this notion.

While the description of the mast cell location in the epidermal skin layer in amphibians is novel, it is not known how representative these results are across species ranging in chytridiomycosis susceptibility. No management applications are provided such as methods to increase this defense without the use of recombinant stem cell factor, and more discussion is needed on how the mast cell component (abundance, distribution in the skin) of the epidermis develops or is regulated.

We thank the reviewer for this suggestion. Please find that we have added a paragraph to our revised manuscripts to address possible source(s) of skin mast cells and a statement acknowledging that greater understanding of mast cell biology across distinct amphibian species may be used to develop future strategies for management of amphibian diseases.

We are very thankful to the reviewer for this excellent suggestion but would like to point out that the work presented in our manuscript was driven by comparative immunology questions more than by conservation biology. As such and considering just how little is known about mast cells outside of mammals; we chose not to speculate too much into possible utilities of altering amphibian skin mast cell composition and instead to focus our discussion on the immediate takeaways of the work presented by our paper.

References

(1) Flajnik, M.F. A cold-blooded view of adaptive immunity. Nat Rev Immunol 18, 438-453 (2018).

(2) Mulero, I., Sepulcre, M.P., Meseguer, J., Garcia-Ayala, A. & Mulero, V. Histamine is stored in mast cells of most evolutionarily advanced fish and regulates the fish inflammatory response. Proc Natl Acad Sci U S A 104, 19434-19439 (2007).

(3) Reite, O.B. A phylogenetical approach to the functional significance of tissue mast cell histamine. Nature 206, 1334-1336 (1965).

(4) Reite, O.B. Comparative physiology of histamine. Physiol Rev 52, 778-819 (1972).

(5) Takaya, K., Fujita, T. & Endo, K. Mast cells free of histamine in Rana catasbiana. Nature 215, 776-777 (1967).

(6) Galli, S.J. New insights into "the riddle of the mast cells": microenvironmental regulation of mast cell development and phenotypic heterogeneity. Lab Invest 62, 5-33 (1990).

(7) Babina, M., Guhl, S., Artuc, M. & Zuberbier, T. IL-4 and human skin mast cells revisited: reinforcement of a pro-allergic phenotype upon prolonged exposure. Archives of dermatological research 308, 665-670 (2016).

(8) Hermans, M.A.W. et al. Human Mast Cell Line HMC1 Expresses Functional Mas-Related G-Protein Coupled Receptor 2. Front Immunol 12, 625284 (2021).

(9) Buchan, K.D. et al. A transgenic zebrafish line for in vivo visualisation of neutrophil myeloperoxidase. PLoS One 14, e0215592 (2019).

(10) Aponte-Lopez, A., Enciso, J., Munoz-Cruz, S. & Fuentes-Panana, E.M. An In Vitro Model of Mast Cell Recruitment and Activation by Breast Cancer Cells Supports Anti-Tumoral Responses. Int J Mol Sci 21 (2020).

(11) Jamur, M.C. et al. Mast cell repopulation of the peritoneal cavity: contribution of mast cell progenitors versus bone marrow derived committed mast cell precursors. BMC Immunol 11, 32 (2010).

(12) Walke, J.B. & Belden, L.K. Harnessing the Microbiome to Prevent Fungal Infections: Lessons from Amphibians. PLoS Pathog 12, e1005796 (2016).

Reviewer #2: (Recommendations For The Authors):

We thank the reviewer for their excellent suggestions, their time reviewing this work and their help with this manuscript.

While we were not able to incorporate some of these changes, please find that we have significantly altered our manuscript in accordance with the reviewer’s suggestions from their public review. We feel that we have substantially altered our paper, including providing considerable additional data, supporting the key findings therein.

(1) The heatmap in Figure 1I appears to be scaled data, similar to Figure 4A, in which case the indicated scale numbers are not correct (e.g. they should be -2 to 2, or -3 to 3)

Thank you for the suggestion. Please find that we have changed this figure accordingly.

(2) For Figure 1, additional curated gene lists might better illustrate the difference in cell types, e.g. include the data for a panel of mast cell genes in a heatmap (mcpt1, tpsab1, etc.) and another panel of curated neutrophil genes (e.g. lyz) in a heatmap. If the authors still have leftover RNA, qPCR verification of some of the critical genes (e.g. kit) would add to the rigor of the analysis, as this study is the foundation of a new method for culturing amphibian mast cells.

We thank the reviewer for this suggestion. Unfortunately, we do not have leftover RNA/cDNA and we have not been able to locate mcpt1 or tpsab1 in our DEGs. We anticipate that this issue may stem from the suboptimal annotation of the Xenopus laevis genome. We agree that curating more mast cell/neutrophil genes would be ideal but feel that we have adequately highlighted those genes that are differentially expressed between the two populations in our analysis.

(3) The presentation of counts in Figure 2 is a bit hard to interpret. Although it is mentioned that everything is statistically significant, explicitly showing statistics for each gene would be better. One possibility would be to use a volcano plot (p-value vs log2 fold change) and highlight the genes shown in Figure 2, potentially with an accompanying heat map to show replicate variability.

We thank the reviewer for this suggestion. We entertained presenting the data as volcano plots or heat maps, but in the end felt that the bar graphs better conveyed the information that we are hoping to get across. Please note that the error bars in the bar graph depict the replicate variability. Please also note that to highlight that all the depicted genes were differentially expressed, we italicized the statement in the corresponding figure legend: “All depicted genes were significantly differentially expressed between the two populations”.

(4) Narratively, it might make more sense to put Figure 4A-C with Figure 3.

We thank the reviewer for this suggestion. Please find that we significantly revised most of our figures to better convey the content therein. We combined the content of Figure 4A-C with Figure 5A-C and added data on epidermal thickness under different conditions into this figure; Figure 5 of our revised manuscript.

(5) If possible, complementing the skin RNA-seq from rSCF treatment in Bd infection with skin RNA-seq from rCSF3 treatment to compare effects on transcriptional programs of barrier function, etc would elevate this study and add additional insights into cutaneous inflammation in the setting of Bd infection.

We thank the reviewer for this suggestion. We anticipate that the skin inflammation caused by Bd infection is not due solely to neutrophil infiltration and artificially altering the frog skin neutrophil content would thus not recapitulate chytridiomycosis progression. We completely agree that it would be valuable to examine barrier functions in control and mast cell-enriched, Bd-infected frogs. This is something that we hope to pursue further in future studies but feel that together with our additional findings, we are presenting a significant amount of data to constitute a stand-alone story.

(6) In Figure S1A, analyzing only 3 AMP genes by qPCR is perhaps too focused. As a control, it would be useful to also test some genes known to be functionally important in neutrophil anti-microbial responses, e.g. lyz. Expanding on this experiment by performing RNA-seq on Bd-treated, bone-marrow-derived mast cells and neutrophils would be a great addition to the manuscript and an important resource for future studies in the field. The fact that the use of rSCF (or rCSF3) enables the differentiation of these cells in large numbers of pure populations presents this unique opportunity. Although IL-4 did not end up affecting mucus production, clues to the mediator(s) of this mast cell-dependent effect may be found with unbiased RNA-seq after exposure to Bd.

We thank the reviewer for this suggestion but would like to point out that our manuscript is focused on mast cells rather than neutrophils. We also believe that in vitro exposure of leukocytes to Bd is not the most physiologically relevant model of what would happen to skin-resident and incoming immune cell subsets, since Bd primarily infects top-most keratinocytes. We anticipate that rather than coming into direct contact with the fungus, cells like mast cells and neutrophils are responding to Bd-produced and infected cell-produced products. For this reason, we did not perform RNA-seq analysis of in vitro derived mast cells or neutrophils stimulated with Bd. As we develop more X. laevis-specific reagents, we hope to revisit the question of infected skin mast cell and neutrophil gene expression profiles but are not in a position to ask these questions at this time.

This work is also guided by a finite budget, and we feel that together with our significant additional findings described in our revised manuscript, we are presenting a substantial amount of work to constitute a stand-alone story and manuscript.

Reviewer #3 (Recommendations For The Authors):

The following are minor edits needed in the text and figure legends:

Standardize terms such as IL4 instead of il4 or ril4 vs rIL4 throughout. Also, r-SCF vs rSCF.

Thank you. Please find that we have standardized such terms throughout our revised manuscript. Please note that we are adhering to the convention that gene names are in lower case, protein names are in upper case and recombinant protein names are preceded by an ‘r’.

Pg 9 Change "In contract" to "In contrast".

Thank you and changed accordingly.

Fig 4 - Perhaps indicate if results in addition to 7dpi are also available.

Please find that we analyzed Bd loads in control and mast cell-enriched, infected frogs after 10 dpi. This data is presented in Figures 3 and 4 of our revised manuscript.

Similarly in Fig. 5, are results other than 10dpi available in the supplement?

Please find that the results from the microbiome studies are presented in supplemental figure 3 (Fig. S3). Please note that the results presented in original manuscript Fig. 5A-C - revised manuscript Fig. 5B-E depict data for 21 dpi, which is the longest examined infection timepoint. We present data from 1 and 10 dpi in Fig. 4 of our revised manuscript.

Indicate why these days were chosen in the methods.

Please find that we indicated why the experimental timepoints were chosen, in the methods section of our revised manuscript.

Fig S1 legend has errors in describing which panels are for which asterisks.

Fig. S3 legend indicates panels F and G.

Thank you. Please find that we revised our supplemental figures and amended the corresponding figure legends.

Version 2

Abstract

Abstract

Global amphibian declines are compounded by deadly disease outbreaks caused by the chytrid fungus, Batrachochytrium dendrobatidis (Bd). Much has been learned about the roles of amphibian skin-produced antimicrobial components and microbiomes in controlling Bd, yet almost nothing is known about the roles of skin-resident immune cells in anti-Bd defenses. Mammalian mast cells reside within and serve as key immune sentinels in barrier tissues like skin. Accordingly, we investigated the roles of Xenopus laevis frog mast cells during Bd infections. Our findings indicate that enrichment of X. laevis skin mast cells confers significant anti-Bd protection and ameliorates the inflammation-associated skin damage caused by Bd infection. This includes a significant reduction in Bd-infected skin infiltration by neutrophils. Augmenting frog skin mast cells promotes greater mucin content within cutaneous mucus glands and protects frogs from Bd-mediated changes to their skin microbiomes. Mammalian mast cells are known for their production of the pleiotropic interleukin-4 (IL4) cytokine and our findings suggest that the frog IL4 plays a key role in conferring the effects seen following frog skin mast cell enrichment. Together, this work underlines the importance of amphibian skin-resident immune cells in anti-Bd defenses and illuminates a novel avenue for investigating amphibian host-chytrid pathogen interactions.

Introduction

Catastrophic declines of hundreds of amphibian species across six continents have been causally linked to the chytrid fungi, Batrachochytrium dendrobatidis (Bd) and Batrachochytrium salamandrivorans (Bsal)1, 2. Motile Bd zoospores readily colonize keratinized skin of adult amphibians and keratinized mouthparts of tadpoles3. Bd skin infections culminate in chytridiomycosis, ultimately disrupting the function of this respiratory and barrier tissue3, 4, 5. Effectively combatting chytrid infections requires a holistic understanding of amphibian cutaneous immune defenses. Research efforts up to this point have focused on antifungal capacities of amphibian skin-produced antimicrobial peptides6, 7; commensal antifungal products7, 8, 9; antifungal properties of mucus9, alkaloids10, and lysozymes11, 12 as well as the roles of antibodies4,13. Many of these studies suggest that Bd exposure can elicit some immune protection14. However, the contribution of skin-resident immune cells to amphibian anti-Bd defenses remains almost entirely unexplored.

Mammalian mast cells serve as sentinels of mucosal and connective tissues, including in barrier tissues like skin, where they maintain homeostasis and regulate immune responses15. Other granulocyte-lineage cells such as neutrophils are generally not found in healthy tissues and only extravasate into sites of inflammation16. Consequently, mast cells are among the first immune cells to recognize and respond to skin-infiltrating pathogens. When activated, mast cells release pre-formed and de novo-generated immunomodulatory compounds that may serve to elicit, exacerbate, or ameliorate inflammatory responses15. One of these mast cell-produced mediators, the interleukin-4 (IL4) cytokine dampens inflammation and promotes tissue repair17. Cells bearing hallmark mast cell cytology have been reported across a range of non-mammalian species18, 19, including amphibians20. Notably, the principal mast cell growth factor, stem cell factor (SCF, KIT ligand) required for mast cell differentiation and survival21 is expressed by all vertebrates examined to-date.

Here, we combine comprehensive in vitro and in vivo approaches to define the roles of amphibian (Xenopus laevis) mast cells during Bd infections. Our results provide compelling evidence that skin-resident immune cells contribute to anti-Bd defenses.

Results

Frog mast cells possess archetypal mast cell cytology and transcriptional profiles

We produced X. laevis recombinant (r)SCF, and used this reagent to generate mast cell cultures from bone marrow-derived myeloid precursors22. Mast cells were compared to bone marrow-derived neutrophilic granulocytes (hereafter referred to as ‘neutrophils’), differentiated using a recombinant X. laevis colony-stimulating factor-322 (rCSF3, i.e., granulocyte colony-stimulating factor; GCSF). While the neutrophil cultures were comprised of cells with hyper-segmented nuclei and neutral-staining cytoplasms (Fig. 1A), the mast cell cultures consisted predominantly of mononuclear cells with basophilic cytoplasm (Fig. 1B). We confirmed the granulocyte-lineage of X. laevis mast cells using specific esterase (SE) staining (Fig. 1D). As expected, X. laevis neutrophils were also SE-positive (Fig. 1C). Mast cell and neutrophil morphology was further explored with electron microscopy (Fig. 1E-H). SEM imaging demonstrated that X. laevis mast cells possess extensive folding of their plasma membranes (Fig. 1F). This mast cell-characteristic membrane ruffling appeared as projections resembling pseudopods via TEM, which further revealed electron-dense heterogenous granules, few mitochondria, and round to elongated nuclei (Fig. 1H) typical of mammalian mast cells23. X. laevis neutrophils also exhibited pronounced membrane ruffling (Fig. 1E) but strikingly distinct intracellular appearance including multilobed nuclei (Fig. 1G).

X. laevis bone marrow-derived mast cells possess classical mast cell cytology and transcriptional profiles.

Neutrophils (A, C, E, G) and mast cells (B, D, F, H) were stained with Giemsa (A, B) and Leder to visualize specific esterase activity (SE) (C, D) or imaged with scanning and transmission electron microscopy (SEM: E, F and TEM: G, H). (I) Heat map of the top 30 differentially expressed genes (DEGs) identified with RNA sequencing analyses of X. laevis mast cell (N=4) and neutrophil (N=4) cultures. Log2fold change in expression represented as color scale.

Frog mast cells and neutrophils exhibited distinct transcriptional profiles of immune-related genes including those encoding lineage-specific transcription factors, immune receptors, downstream signaling components and adhesion molecules, as well as non-immune genes (Fig. 1I). Frog mast cells and neutrophils each expressed greater levels of lineage-specific transcription factors associated with mammalian mast cell (gata1, gata2, and mitf)24 and neutrophil (cebp family members)25 counterparts, respectively (Fig. 2A). Notably, mast cells expressed greater levels of enzyme and cytokine genes associated with tissue remodeling (carboxypeptidase-3; cpa326), immune suppression (indoleamine 2,3 dioxygenase-1; ido127) and amelioration of cutaneous inflammation (leukemia inhibitory factor, lif28; Fig. 2B). Conversely, neutrophils expressed predominantly proinflammatory enzymes and cytokine genes such as leukotriene 4 hydrolase (lta4h; Fig. 2B) and tumor necrosis factor alpha (tnfa, Fig. 2B). In addition, mast cells and neutrophils each had greater expression of genes encoding their respective growth factor receptors, kit and csf3r (Fig. 2B).

Frog mast cells and neutrophils possess gene profiles similar to their mammalian counterparts.

The differentially expressed genes from the RNA sequencing analyses of X. laevis mast cells and neutrophil cultures were profiled for those encoding (A) transcription factors associated with mast cell- or neutrophil-specific lineages and (B) granulocyte antimicrobial components and growth factor receptor genes. All depicted genes were significantly differentially expressed between the two populations, N=4 per group.

Enriching mast cells in frog skin offers protection against Bd

Although all granulocyte-lineage cells possess SE activity, mast cells are the predominant mononuclear granulocytes to reside in vertebrate tissues29. Therefore, we selectively enriched mast cells in X. laevis skin via subcutaneous rSCF administration (note SE-stained cells indicated by arrows in r-ctrl-injected skins, Fig. 3A, versus r-SCF-injected skins, Fig 3B). We confirmed SE-positive cells in rSCF-treated skins also possessed round-oval nuclei (Fig. 3C). Maximum mast cell enrichment was observed 12 hours post injection (hpi) of rSCF (Fig. 3D).

Enriching frog cutaneous mast cells lowers Bd loads.

Representative images of SE stained (A) control and (B) mast cell-enriched skin 12 hpi. (C) We confirmed the enriched population was composed of mono-morphonuclear cells. (D) Mast cell enrichment was optimized across several time points by quantifying SE-positive cells per field of view under 40x magnification. Results represent means ± SEM from 3 animals per time point (2 experimental repeats). (E) Mast cell-enriched and control dorsal skins were collected from X. laevis 10- and 21-dpi. Bd loads are represented as the number of zoospore genomic equivalents (GE) × 105 per μl of total input DNA. Time points were analyzed independently. Results represent means ± SEM from 7 animals per experimental group (N=7). Asterisks indicate significance: p < 0.05 by (D) one-way ANOVA with Tukey post-hoc analysis or (E) Student’s t-test.

We next examined the consequences of enriching frog skin mast cells on Batrachochytrium dendrobatidis (Bd) infection outcomes. To this end, X. laevis were subcutaneously administered with rSCF or the r-ctrl, infected with Bd, and the skin fungal loads assessed at 10- and 21-days post infection (dpi). At 10 dpi, skin mast cell-enriched X. laevis possessed significantly lower Bd loads than r-ctrl administered animals (Fig. 3E). By 21 dpi, both r-ctrl and rSCF-administered groups possessed substantially greater Bd loads, although the mast cell-enriched animals continued to show significantly lower skin fungal loads (Fig. 3E).

Mammalian mast cells may be labeled in situ by using avidin to visualize the heparin-containing mast cell granules30. Using this approach, we confirmed that frog bone marrow-derived mast cells likewise possess heparin-containing granules (Fig. 4A&B). We next used this staining approach to visualize and enumerate mast cells in the skins of control, mast cell-enriched and Bd-challenged animals (Fig. 4C-I). We observed that frog skin mast cell enrichment progressively waned with time in mock-infected animals, and by 21 days post mock infection these animals had skin mast cell numbers comparable to r-ctrl-injected, mock-infected control animals (Fig. 4C, D). Conversely, compared to mock-infected control animals, the rSCF-injected, Bd-challenged frogs maintained significantly greater mast cell numbers in their skin throughout the 21-day infection study (Fig. 4C, E-H). Interestingly, r-ctrl-injected, Bd-infected animals exhibited significantly increased skin mast cell numbers at 10- and 21-days of infection, compared to uninfected control frogs (Fig. 4C, I). Most of these mast cells were observed in the epidermal layer, spreading out between the skin epithelial cells (Fig. 4G).

Heparin content and skin localization of frog mast cells.

Frogs were administered r-ctrl or rSCF subcutaneously, mock or Bd-challenged and skin examined after 1, 3, 10 and 21 dpi. (A, B) Representative images (cultures derived from 5 individual frogs) of bone marrow-derived frog mast cells, stained with fluorescently labeled avidin to visualize the heparin-containing mast cells (avidin: red; nuclei: blue; actin: green). (C-I) Skin tissue from control (r-ctrl) or mast cell-enriched (rSCF), mock-(not shown) and Bd-infected (D-I) X. laevis were stained with fluorescently labeled avidin to visualize mast cells therein (N=6 animals per treatment group). Images were inverted in ImageJ for greater contrast and visibility. (C) Heparin-containing mast cells were enumerated and depicted as means + SEM of heparin-positive cells per field of view, N=6 animals per treatment group. Asterisks indicate statistical significance from r-ctrl: p < 0.05. Representative images of heparin-containing mast cells in the skins of (D) r-ctrl animals 1 dpi; (E) rSCF-administered frogs 1 dpi; (F, G) rSCF-administered frogs 10 dpi; (H) rSCF-administered frogs 21 dpi and (I) r-ctrl-treated animals 21 dpi with Bd.

Mast cells protect frogs from Bd-elicited inflammation-associated pathology

To explore potential mechanisms of mast cell-mediated protection against Bd, we compared the gene expression profiles of r-ctrl- and rSCF-administered, Bd-infected frog skins at 21 dpi. Among the top differentially expressed genes, we noted mast cell-enriched, Bd-infected skins possessed greater transcripts for genes associated with cutaneous strength and integrity (lamc2)31, suppression of cell migration (b3gnt3.1)32, as well as ion and nutrient flow (gjb3l)33 (Fig. 5A). Moreover, mast cell-enriched Bd-challenged skin exhibited greater expression of genes associated with protection of the mucosa and epithelial healing (ttf3.6s)34 and mucus production (duoxa1.s35, gabrp36; Fig. 5A). In striking contrast, skins from control Bd-infected frogs revealed greater expression of genes associated with leukocyte infiltration and inflammation (e.g., ccl19, cxcl16, adamts13, csf3r; Fig. 5A). These transcriptional profiles were supported by our histological observations wherein control Bd-infected skins exhibited hyperkeratosis, epidermal hyperplasia, jagged stratum corneum, and extensive leukocyte infiltration (Fig. 5B), while mast cell-enriched Bd-infected tissues appeared considerably less afflicted by these pathologies (Fig. 5C). Quantification of Bd-infected skin epidermal thickness confirmed that mast cell-enriched animals possessed significantly less thickened epidermal skin compared to control (r-ctrl), Bd-infected animals (Fig. 5D).

Consequences of cutaneous mast cell enrichment.

(A) RNAseq analysis of skin tissue from control (r-ctrl) or mast cell-enriched (rSCF) Bd-infected X. laevis at 21 dpi. Heat map of the top 30 DEGs, numbers matched to colors represent log2 fold change in expression, N = 4 r-ctrl-treated, Bd-infected and 4 rSCF-treated, Bd-infected skin samples. (B & C) Representative images of control and mast cell-enriched, Bd-infected skins, 21 dpi, demonstrating differences in epidermal thickening and mucus gland filling. Mucin content was visualized in cutaneous mucus glands with Alcian Blue/PAS stain. Mucus glands are denoted by ‘m’, and epithelia are denoted by ‘Ep’. ImageJ software was used to determine the means + SEM of (D) skin epithelial thickness and (E) percent mucus gland filling (N=6). Asterisks indicate significance: p < 0.05 by one-way ANOVA with Tukey post-hoc analysis.

Because mast cell-enriched frog skins had greater expression of genes associated with mucosal tissue integrity and mucus production (Fig. 5A), we also investigated whether the anti-Bd protection identified in mast cell-enriched skins could be due, at least in part, to differences in mucus production. Interestingly, the skin mucus glands of mast cell-enriched, mock- and Bd-infected frogs were significantly more filled than those of mock- and Bd-infected control animals (Fig. 5B,C,E).

Cutaneous neutrophil enrichment results in increased Bd fungal loads

Neutrophils are one of the first leukocytes to infiltrate infected tissues, typically amplifying inflammation37. All vertebrate neutrophils depend on CSF3 for their differentiation and function38, and our previous work has demonstrated the gene encoding the CSF3 receptor (csf3r) is a marker of X. laevis neutrophils39, 40. Because csf3r expression was markedly elevated in control over mast cell-enriched skins of infected frogs (Fig. 5A), we examined the neutrophil content in the skins of these animals over the 21-day course of Bd infection (Fig. 6A-E). To this end, we used in situ hybridization analysis of skin-myeloperoxidase, a marker of neutrophils41. While the skins of r-ctrl- and rSCF-administered, mock-infected frogs contained relatively few neutrophils (Fig. 6A, E), the skins of r-ctrl-injected, Bd-infected animals had significantly elevated neutrophil numbers (Fig. 6B,C,E). Conversely, at both examined timepoints (10, 21 dpi), mast cell-enriched frog skins possessed neutrophil levels similar to those seen in the uninfected animals (Fig. 6D,E).

Consequences of cutaneous neutrophil enrichment.

(A-E) Frogs were administered with r-ctrl or rSCF, mock- or Bd-infected and examined in situ for neutrophil content via by RNAScope analyses of myeloperoxidase (mpo) transcripts. Representative images of (A) r-ctrl-injected frog skins 10 days post mock infection; (B) r-ctrl-injected frog skins 10 days post Bd infection; (C) r-ctrl-injected frog skins 21 days post Bd infection; and (D) r-SCF-administered frog skins 21 days post Bd infection. (E) Means + SEM of mpo-positive neutrophils per field of view of r-ctrl- and rSCF-administered, mock- or Bd-challenged frog skins 10 or 21 dpi, (N=6). Asterisks indicate significance from control: p < 0.05 by one-way ANOVA with Tukey post-hoc analysis. Representative specific-esterase staining of skin tissues from frogs were administered with (F) r-ctrl or (G) rCSF3, N=4. Arrows denote specific esterase-positive cells. Ep: epithelium. (H) Bd loads in control and neutrophil-enriched skin tissues 7 dpi, N=6. Asterisks indicate significance: p < 0.05 by one-way ANOVA with Tukey post-hoc analysis.

We next assessed the consequence of enriching frog skins for neutrophils via subcutaneous rCSF3 administration. We confirmed neutrophil enrichment peaked 12 hp rCSF3 injection (Fig. S1) and resulted in a thickened epidermis in comparison with r-ctrl injected skins of otherwise healthy animals (i.e., no Bd; Fig. 6F, G). When challenged with Bd, frogs with neutrophil-enriched skin possessed significantly greater Bd loads than control frogs (Fig. 6H). This suggests that neutrophil-mediated inflammation may be exacerbating Bd infections.

Enrichment of frog skin mast cells alters skin microbial composition

Amphibian skin-produced mucus may offer antimicrobial protection and serve as a selective substratum for commensal microbes, many of which are antifungal9. We found no significant differences in direct Bd-killing capacities of mucus isolated from mock- or Bd-challenged control or mast cell-enriched frogs (Fig. S2). By contrast, we observed substantial differences in skin microbiomes, including changes in bacterial composition and richness as well as relative abundances of Bd-inhibitory bacteria (Fig. 7, Fig. S3). A total of 1645 bacterial amplicon sequence variants (ASVs) were identified from 20 bacterial phyla, seven of which were predominant (Fig. 7A). Of these, Verrucomicrobiota were only present on uninfected animals, whereas Acidobacteriota was only seen after 21 dpi on both control and mast cell-enriched, infected animals (Fig. 7A).

Cutaneous mast cells protect skin microbial communities.

Control (r-ctrl-injected) or mast cell-enriched (rSCF-injected) X. laevis were mock-infected or challenged with Bd for 21 days. (A) Microbial phyla distribution across groups. Low abundance phyla (< 5% relative abundance are not shown). At 10 dpi (B) community composition (Jaccard distances shown with 80% confidence ellipses) differed among all treatments. (C) Relative abundance of Bd-inhibitory bacteria and (D) bacterial richness were examined in control and mast cell enriched frogs, 10 days post Bd or mock challenge. Letters above bars indicate statistically different groups.

At 10 dpi, mast cell-enrichment resulted in a nominal shift in community composition compared to control frogs (Fig. 7B). Notably, while control, Bd-infected animals exhibited a drastic shift in community composition, mast cell-enriched animals possessed substantially less deviated community composition (Fig. 7B), suggesting that these cells are somehow counteracting the adverse effects of Bd on microbiome structure. These mast cell-mediated effects persisted to 21 dpi (Fig. S3A).

At 10 dpi, mast cell-enriched and mock-infected frogs possessed significantly greater abundance of Bd-inhibitory bacteria such as Chyrseobacterium sp., compared to control, mock infected animals (Fig. 7C)42. This suggests that mast cells may promote skin flora composition that is more antifungal. Control (non-enriched) Bd-infected frogs possessed significantly greater abundance of Bd-inhibitory bacteria than all other treatment groups (Fig. 7C). While mast cell-enriched, Bd-infected frogs had lower abundance of Bd-inhibitory bacteria than control infected frogs, they possessed higher abundance of inhibitory taxa than uninfected control animals (Fig. 7C). The Bd-inhibitory bacteria seen in greater abundance on mast cell-enriched, Bd-infected animals, included Roseateles sp., Flavobacterium sp., and Kaistia sp. We did not see significant differences in Bd-inhibitory bacteria across the treatment groups at 21 dpi (Fig. S3B).

Mast cell-enriched uninfected frogs exhibited increased frog skin bacterial richness at 10 dpi (Fig. 7D). While control Bd-infected animals exhibited significantly reduced skin microbial richness, mast cell-enriched Bd-infected frogs did not exhibit such a reduction in bacterial richness (Fig. 7D), supporting the idea that mast cells may be counteracting the adverse effects of Bd on skin microbiome composition.

Frog mast cells alter skin antimicrobial peptide gene expression

Amphibians rely heavily on skin-produced antimicrobial peptides (AMPs) for antifungal protection43, and mast cells produce antimicrobial AMPs44. We thus examined whether mast cells could be sources of such AMPs during Bd infections. As anticipated, Bd-challenged mast cells, but not Bd-challenged neutrophils upregulated their expression of the AMP-encoding genes, pgla and magainin (mag, Fig. 8A).

The roles of IL4 in mast cell-mediated skin anti-Bd protection.

(A) Mast cells and neutrophils derived from bone marrow of 6 individual frogs (N=6) were co-cultured with Bd (5 fungal cells per granulocyte) for 6 hrs prior to gene expression analyses of the antimicrobial peptide genes PGLa (pgla) and magainin (mag) or interleukin-4 (il4). (B) Il4 gene expression in skins of control and mast cell-enriched, mock- or Bd-infected animals, N=6. Representative images of frogs administered with (C) r-ctrl or (D) rIL4 and infected with Bd for 21 days, N=7. Ep: epidermis; m: mucus gland. Means + SEM of (E) percent mucus gland filling, (F) skin Bd loads, (G) epidermal thickness and (H) mpo-positive neutrophils, per field of view of r-ctrl- and rSCF-administered, mock- or Bd-challenged frog skins 21 dpi, (N=7). Representative images of mpo-positive neutrophils in (I) control and (J) rIL4-treated frog skins, 21 dpi. Ep: epidermis; arrows: mpo-positive neutrophils. Asterisks indicate significance: p < 0.05 by one-way ANOVA with Tukey post-hoc analysis.

To follow up these observations in vivo, we examined pgla and mag gene expression in the skins of mast cell-enriched and Bd-challenged animals. Compared to control (r-ctrl) animals, mast cell-enriched frogs did not have elevated mRNA levels of mag or pgla after 12 hrs of rSCF administration (Fig. S4A&B). After 10 days of mock-infection, mast cell-enriched animals possessed significantly greater skin expression of mag and pgla than control animals (Fig. S4A&B). We did not see significant differences in skin mag or pgla gene expression between Bd-challenged control and mast cell-enriched frogs after 10 dpi (Fig. S4). By 21 days of challenge, the rSCF-administered mock-infected frogs possessed lower gene expression levels of both AMPs (significantly so for pgla, Fig. S4A&B), possibly due to some sort of compensatory effect. At 21 dpi, mast cell-enriched frog skins had greater (albeit not significantly so) pgla expression than control frogs (Fig. S4B).

Frog mast cell-expressed interleukin-4 confers anti-Bd protection

Mammalian mast cells are recognized as potent producers of the pleotropic anti-inflammatory cytokine, interleukin-4 (il4)45. Notably, X. laevis mast cells challenged in vitro with Bd significantly upregulated their il4 gene expression, whereas almost no il4 expression was detected from either unstimulated or Bd-challenged neutrophils (Fig. 8A).

When we examined the expression of il4 in the skin of control and mast cell-enriched animals, we found that 12 hours following rSCF administration (time of Bd challenge), the mast cell enriched frog skins possessed significantly greater transcripts of il4 than control animals (Fig. 8B). By 10 and 21 dpi, both control and mast cell-enriched mock-infected animals possessed baseline skin il4 expression (Fig. 8B). At 10 days of Bd infection, control animals possessed increased (not significantly) skin il4 gene expression whereas mast cell-enriched frog skins had significantly elevated il4 transcript levels (Fig. 8B). By 21 dpi, Bd-infected control animals had baseline skin il4 expression and the mast cell enriched animals had elevated, albeit not significant expression of this cytokine gene in their skins (Fig. 8B).

To examine whether IL4 could confer the anti-Bd effects seen following frog skin mast cell enrichment, we produced this X. laevis cytokine in recombinant form (rIL4) and confirmed that subcutaneous injection of rIL4 augmented expression of genes typically activated by the mammalian IL446 (cd36, metalloproteinase inhibitor 3-timp3, and monoamine oxidase A-maoa; Fig. S5A).

We next administered frogs subcutaneously with rIL4 or r-ctrl, challenged them with Bd and examined the key parameters affected by skin mast cell-enrichment, including skin mucus gland filling, skin Bd loads, and reduction in Bd-associated skin inflammation. When we examined the mucus content of control and rIL4-administered animals, we found that frogs treated with rIL4 possessed greater skin mucus gland filling than control animals and irrespective of mock- or Bd-challenge (Fig. 8C-E). At 21 dpi (but not at 10 dpi), rIL4-administered frogs possessed significantly lower skin Bd loads (Fig. 8F) and exhibited significantly less epidermal thickening (Fig. 8C,D,G) and less neutrophil infiltration (Fig. 8H-J) than control animals. Conversely, subcutaneous administration of rIL4 to animals with active Bd-infection did not alter their fungal loads (Fig. S5B), suggesting a critical point during which IL4 may offer protection against this fungal pathogen.

Discussion

Amphibian extinction rates far outpace those of any other vertebrate class47. It is now well-established that chytrid fungi are major contributors to these declines, and strikingly, are considered the greatest infectious disease threat to biodivsersity48. The development of effective mitigation strategies though, is hindered by incomplete understanding of amphibian immune defenses and skin integrity. In this respect, while mast cells are recognized as key immune sentinels of tissues such as skin49, relatively little is known about this cell lineage outside of mammals. Our findings provide the first in-depth functional analyses of these cells in amphibians and explore their protective mechanisms during chytrid infections. This work presents a unique perspective on the evolution of mast cell functionality and will serve as a new avenue to explore ways to counteract the amphibian declines.

Mammalian mast cells are potent producers of the pleotropic anti-inflammatory interleukin-4 (IL4)45 cytokine and our findings suggest that this is also true of frog mast cells. In mammals, IL4 also plays roles in tissue repair17 and mucus production by intestinal goblet cells50, which aids in the ‘weep and sweep’ pathogen elimination and may contribute to the maintenance of commensal communities in the gut51. This is consistent with our findings, which suggest that IL4 may be a central means by which frog mast cells confer protection against Bd, by counteracting Bd-elicited inflammation, including minimizing neutrophil infiltration, maintaining skin integrity, and promoting mucus production by skin mucus glands. The outcomes of cytokine signaling are critically dependent on timing, locations, and context50. This possibly explains why administering rIL4 prior to Bd challenge results in significant anti-Bd protection whereas treating already infected animals with this cytokine did not reduce fungal loads.

We anticipate that in addition to IL4 production, frog mast cells confer the observed antifungal protection through a myriad of additional mechanisms. Indeed, we observed elevated expression of antimicrobial peptide genes in the skins of mock-infected mast cell-enriched frogs 10 days after rSCF administration. The fact that we did not see the same expression patterns following Bd challenge could potentially reflect the highly immunomodulatory capacities of this pathogen52. Undoubtedly, frog skin-resident mast cells have evolved to contribute to and modulate skin antifungal antimicrobial peptide production just as Bd has undoubtedly coevolved to target this aspect of amphibian skin defenses.

Bd infections caused major reductions in bacterial taxa richness, changes in composition and substantial increases in the relative abundance of Bd-inhibitory bacteria early in the infection. Similar changes to microbiome structure occur during experimental Bd infections, such as in red-backed salamanders and mountain yellow-legged frogs8, 53. In turn, progressing Bd infections corresponded with a return to baseline levels of Bd-inhibitory bacteria abundance and rebounding microbial richness, albeit with dissimilar communities to those seen in control animals. These temporal changes indicate that amphibian microbiomes are dynamic, as are the effects of Bd infections on them. Indeed, Bd infections may have long-lasting impacts on amphibian microbiomes8, including the possible establishment of ‘ecological memory’, resulting in greater adaptation to subsequent pathogen exposures9, 54. While Bd infections manifested in these considerable changes to frog skin microbiome structure, mast cell enrichment appeared to counteract the deleterious effects to their microbial composition. It will thus be interesting to learn from future studies how frog mast cells impact skin microbiome ecological memory. Presumably, the greater skin mucosal integrity and mucus production observed after mast cell enrichment served to stabilize the cutaneous environment during Bd infections, thereby ameliorating the Bd-mediated microbiome changes. While this work explored the changes in established antifungal flora, we anticipate the mast cell-mediated inhibition of Bd may be due to additional, yet unidentified bacterial, viral, or fungal taxa. Intriguingly, while mammalian skin mast cell functionality depends on microbiome elicited SCF production by keratinocytes55, our results indicate that frog skin mast cells in turn impact skin microbiome structure and likely their function. It will be interesting to further explore the interdependent nature of amphibian skin microbiomes and resident mast cells.

In contrast to the mast cell-mediated effects, enrichment of neutrophils seemed to result in greater Bd skin burdens. This suggests that neutrophils may result in a greater level of inflammation, which is not protective in the context of Bd infections. This is consistent with other studies that suggest that a strong immune response in the skin compartment of some species may be detrimental56, 57.

Amphibian skin is much thinner and more permeable than that of mammals7, and as such, arguably represents a more penetrable barrier to pathogens. Because mammalian skin is relatively impermeable, mast cells are absent from healthy mammalian epidermises and are instead found exclusively in their dermal layers58. However, mammalian mast cells may infiltrate the skin epidermal layer during diseases such as dermatitis59. In contrast, we showed here and have observed across several classes of amphibians16 that mast cells are found in both epidermal and dermal layers. This localization has presumably evolved to support the more intimate contact between amphibian skin and their environments. Considering the importance of amphibian cutaneous integrity to their physiology60, it is likely that skin-resident mast cells co-evolved to support skin immunity, physiology, and symbiotic microbiota. Ongoing research continues to reveal functional differences between mammalian connective tissue and mucosa-resident mast cells61, and we suspect that there is similarly much to learn about the distinct physiological and immune roles of amphibian epidermal and dermal mast cells.

Mammalian mast cells are thought to arise from two distinct lineages. Fetal-derived progenitors seed peripheral tissues during the neonatal life62 and mature into long-lived connective tissue mast cells (CTMCs)63. Conversely, inducible mucosal mast cells (MMCs) arise from bone marrow–derived mast cell progenitors (MCps) and are recruited to mucosal tissues in response to inflammation64, 65. The connective tissue mast cells are thought to possess significantly greater levels of heparin than bone marrow-derived mast cells66. Interestingly, the in vitro frog bone marrow-derived mast cells possessed substantially lower heparin content than what we observed in the skin-resident and Bd-recruited mast cells. Unlike the mammalian skin, frog skin is both a connective tissue and a mucosal barrier, begging the question of how the frog skin mast cells compare to the mammalian CTMCs and MMCs. It is interesting to consider that frog skin mast cells observed at steady state, following rSCF-enrichment, and recruited/expanded in response to Bd, all had substantial heparin levels. Possibly, the expansion of frog skin mast cell numbers in response to both rSCF and Bd is facilitated by skin-resident precursors, explaining the high heparin content observed in these populations. Of course, it is equally possible that amphibian mast cell ontogeny and heparin content are not the same as mammals. It will be invaluable to learn with future studies the origins, fates and immune consequences of distinct frog skin mast cell lineages and subsets.

It has become apparent that amphibian host-Bd interactions are highly complex and multifaceted while different amphibian species exhibit marked differences in their susceptibilities to this devastating pathogen2, 4. The findings described here emphasize the importance of skin-resident mast cells for successful anti-Bd defenses and demonstrate that these immune sentinels are intimately linked to many aspects of frog skin physiology. Our results indicate that when mast cells are enriched, the ensuing changes in the skin allow for greater resistance to infection by developing zoospores. Presumably, distinct amphibian species have evolved disparate interconnections between their skin mast cells and their cutaneous defenses, as dictated by their respective physiological and environmental pressures. In turn, these species-specific differences likely dictate whether and to what extent the skin-resident mast cells of a given amphibian species recognize and appropriately respond to Bd infections. We postulate that such differences may contribute to the differences in susceptibilities of amphibian species to pathogens like chytrid fungi. Greater understanding of these relationships across disparate amphibian species holds promise of harnessing this knowledge towards possible development of preventative and therapeutic strategies against infectious diseases like chytridiomycosis.

Methods

Animals

Outbred 1 year-old (1.5-2”), mixed sex X. laevis were purchased from Xenopus 1 (Dexter, MI). All animals were housed and handled under strict laboratory regulations as per GWU IACUC (Approval number 15-024).

Recombinant cytokines and bone marrow granulocyte cultures

The X. laevis rSCF, rCSF3 and rIL4 were generated as previously described for rCSF367 and as detailed in the supplemental materials.

Bone marrow isolation, culture conditions, and establishment of neutrophil cultures have been previously described68. Mast cell cultures were generated according to protocols adapted from Koubourli et al. (2018) and Meurer et al. (2016) 69, 70. Isolated bone marrow cells were treated with 250 ng/µl of rSCF on Day 0, Day 4, Day 7, and collected for further analysis on Day 9. Cell cultures were maintained at 27°C with 5% CO2 in amphibian medium supplemented with 10% fetal bovine serum and 0.25% X. laevis serum. Neutrophil-like granulocytes were generated as above but with 250 ng/µl of rCSF3 on Day 0, Day 3, and collected for further analysis on Day 5. Cell cultures were maintained at 27°C with 5% CO2 in amphibian serum-free medium supplemented with 10% fetal bovine serum, 0.25% X. laevis serum, 10 μg/mL gentamicin (Thermo Fisher Scientific, Waltham, Massachusetts, USA), 100 U/mL penicillin, and 100 μg/mL streptomycin (Gibco, Thermo Fisher Scientific).

Enrichment of skin granulocyte subsets

Animals were subcutaneously injected between the skin and muscle layers with 5 µg/animal of rSCF, rCSF3, or r-ctrl in 5 µL of saline using finely pulled glass needles. Optimal time course and dose for in vivo mast cell and neutrophil enrichment were determined during preliminary experiments.

Recombinant interleukin-4 treatment

The capacity of the recombinant interleukin-4 (rIL4) to induce expression of genes associated with mammalian IL4 responses were assessed by injecting frogs (N=6 per treatment group subcutaneously with rIL4 (5 µg/animal) or r-ctrl in 5 µl of saline. After 6hrs, animals were sacrificed, and skins were isolated for gene expression analyses.

To examine the effect of rIL4 on Bd loads, frogs were infected with Bd by water bath (105 zoospores/mL) as described below and 1 day later injected subcutaneously, dorsally with rIL4 (5 µg/animal) or r-ctrl in 5 µl of saline. After an additional 9 days of infection, animals were sacrificed and their dorsal skin Bd loads examined.

Bd stocks and fungal challenge

Bd isolate JEL 197 was grown in 1% tryptone broth or on 1% tryptone agar plates (Difco Laboratories, Detroit, MI) supplemented with 100 U/mL penicillin and 100 μg/mL streptomycin (Gibco) at 19°C.

In vitro Bd killing was evaluated by incubating live Bd (maturing zoosporangia) with mast cells or neutrophils at ratios of 5:1 or 1:1 Bd cells per granulocyte. Cells were incubated at 27°C for three days before fungal loads were analyzed by absolute qPCR. Experimental groups were compared to pathogen DNA amounts derived from equal quantities of live Bd plated alone in otherwise identical conditions.

For in vivo infection studies, zoospores were harvested by flooding confluent tryptone agar plates with 2 mL sterile A-PBS for 10 minutes. Twelve hours post rSCF, rCSF3, rIL4, or r-ctrl injection, animals were infected with 107 zoospores or mock-infected in 100 mL of water (105 zoospores/mL). After 3 hrs, 400 mL of water was added to each tank. Skins were collected for histology and gene expression analyses on 1, 10, and 21 dpi with 10 and 21 dpi representing intermediate and later timepoints of infection, respectively.

Histology

Leukocyte cytology and cutaneous SE staining has been described67 and is detailed in the supplemental materials. An Alcian Blue/PAS staining kit (Newcomer Supply, Middleton, WI) was used to quantify mucin content. Histology is detailed in the supplemental materials.

Electron Microscopy

Processing and imaging of cells for transmission and scanning electron microscopy (TEM and SEM) was conducted at the GWU Nanofabrication and Imaging Center (GWNIC). For transmission electron microscopy, cells were fixed as monolayers on six-well plates with 2.5% glutaraldehyde and 1% paraformaldehyde in 0.1 M sodium cacodylate buffer for one hour. Cells were treated with 1% osmium tetroxide in 0.1 M sodium cacodylate buffer for 1 hr. Following washes, cells were en bloc stained with 1% uranyl acetate in water overnight at 4°C. Samples were dehydrated through an ethanol series and embedded in epoxy resin using LX112. Inverted Beem capsules were placed into each tissue culture well to create on face blockfaces for sectioning. Resin was cured for 48 hrs at 60°C. The 95 nm sections were post-stained with 1% aqueous uranyl acetate and Reynold’s lead citrate. All imaging was performed at 80 KV in a Talos 200X transmission electron microscope (Thermo Fisher Scientific, Hillsboro, OR).

For SEM, cells were fixed with 2.5% glutaraldehyde / 1% paraformaldehyde in Sodium cacodylate buffer, followed by 1% OsO4, then dehydrated through an ethyl alcohol series. Coverslips were critical point dried and coated with 2 nm iridium. Cells were imaged using a Teneo Scanning Electron Microscope (Thermo Fisher Scientific).

Analyses of immune gene expression and Bd skin loads

These analyses have been described67 and are detailed in supplemental materials.

RNA sequencing

For transcriptomic profiling, bone marrow-derived neutrophil and mast cell cultures were generated as described above and FACS-sorted according to preestablished size and internal complexity parameters to isolate the respective subsets for further analyses. Sorted cells were immediately processed to extract and purify RNA. Flash frozen samples were sent to Azenta Life Sciences for all library preparation, RNA sequencing, and analyses. In short, polyadenylated RNA was isolated using Oligo dT beads. Enriched mRNAs were then fragmented for first and second strand cDNA synthesis. cDNA fragments were end repaired, 5’ phosphorylated, and dA-tailed. Fragments were then ligated to universal adaptors and PCR-amplified. 150-bp paired-end sequencing was performed on an Illumina HiSeq platform.

FastQC was used to evaluate raw data quality. Adaptors sequences and poor-quality nucleotides were removed from reads using Trimmomatic v.0.36. The STAR aligner v.2.55.2b was used to map these reads to the Xenopus_laevis_9_2 reference genome from ENSEMBL. To determine differential gene expression, featureCount (Subread package v.1.5.2) was first used to count unique gene hits, which were then used with DESeq2 to calculate absolute log2-fold change.

Skin microbiome analyses

Towards microbiome studies, frogs were housed individually (N=6/treatment group). At indicated times, frogs were gently rinsed with sterile deionized water to remove transient microbes and gently swabbed 20 times, dorsally. Genomic DNA was extracted from swabs using a PowerSoil Pro kit on a Qiacube HT (QIAGEN, MD). One-step PCR library prep and dual-index paired-end Illumina sequencing was used to sequence the skin microbiome of individual frogs. A ∼380 base pair region in the V3-V5 region of the 16S rRNA gene using the universal primers 515F-Y (GTGYCAGCMGCCGCGGTAA) and 939R (CTTGTGCGGGCCCCCGTCAATTC) was used for amplification. Negative and positive controls (ZymoBIOMICS D6300 & D6305, Zymo, CA) were included in each round of extraction and PCR. Reactions were done in duplicate for each sample, pooled, cleaned with in-house Speed-beads (in a PEG/NaCl buffer), quantified with a Qubit4 (Invitrogen, MA) and pooled into a final library in equimolar proportion. The pooled library was sequenced on two Illumina MiSeq runs (v3 chemistry: 2×300 bp kit) at the Center for Conservation Genomics, Smithsonian National Zoo & Conservation Biology Institute.

Analyses were performed in the R environment version 4.0.3 (R Core Team, 2020) using methods detailed in the supplemental materials.

Statistical analyses

Differences in transcript expression were calculated with one-way or multi-way ANOVAs followed by Tukey post-hoc tests. Student’s t-tests were used to determine differences in Bd loads between treatments only. Statistical differences in mucin content and mucosome Bd-killing were assessed with the two-way ANOVA calculator available online through Statistics Kingdom. For RNA sequencing, p-values were calculated with the Wald test and were adjusted using the Benjamini-Hochberge procedure.